### 2015-03-16 paper/paper.tex
adding paper text

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

\begin{document}

\title{Solving the grand challenge using an algorithmic an distributed approach}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{2} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Amila Suriarachchi\\
       \affaddr{Colorado State University}\\
       \affaddr{Fort Collins}\\
       \affaddr{Colorado}\\
       \email{amilas@cs.colostate.edu}
% 2nd. author
\alignauthor
Shrideep Pallickara\\
       \affaddr{Colorado State University}\\
       \affaddr{Fort Collins}\\
       \affaddr{Colorado}\\
       \email{shrideep@cs.colostate.edu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Processing complex queries on unbounded event streams in real time, is a challenge for many organizational data processing systems. These systems are expected to process data with minimal latency to generate real time events, and high through put to minimize the required hardware. In this regard DEBS Grand Challenge 2015 focus on evaluating two queries (frequent routes and  profitable cells) in real time with minimal latency and high throughput. These queries involve processing windows of thousands of records. Firstly, such processing demands efficient data structures and algorithms to minimize the processing over head. Secondly, system should partition data to process them in parallel to efficient use of hardware.

In this paper, we present a set of data structures which can be used to evaluate two queries with O(log(n)) time complexity and a data partition technique to process events in parallel. Then we evaluate our solution within signal node as well as in distributed setups. We could process frequent routes query with 173 million data set within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning, clark:pct, braams:babel, herlihy:methodology},
conference
proceedings \cite{clark:pct} or books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\epsfig{file=fly.eps}
\caption{A sample black and white graphic (.eps format).}
\end{figure}

\begin{figure}
\centering
\epsfig{file=fly.eps, height=1in, width=1in}
\caption{A sample black and white graphic (.eps format)
that has been resized with the \texttt{epsfig} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

\begin{figure*}
\centering
\epsfig{file=flies.eps}
\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}
and don't forget to end the environment with
{figure*}, not {figure}!

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The acm\_proc\_article-sp document class file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
\balancecolumns
% That's all folks!
\end{document}

### 2015-03-16 paper/query_implementation.tex
adding query implementation

\section{Query Implementation}
In this section, we elaborate how we have implemented the query logic to achieve O(log(n)) time complexity. The 2015 Grand Challenge problem has two queries. First is to evaluate top ten most frequent routes within last 30 minutes and second is to evaluate top ten profitable areas within last 15 minutes. We can use a composite object to keep route details for each route and fare details to each cell to compare values among routes and cells. However, each time window can have thousands of route and cell details. Therefore an efficient data structure is required to add, remove, update and retrieve top ten values from a set of objects for a given time window efficiently. Further since each event is received with route and cell details, above mentioned operations should perform using route and cell details as the key. In order to support all these functions with O(log(n)) time complexity we developed a data structure called NodeList. Following parts of this section describes how we have achieve this time complexity with \textit{NodeList} data structure and how we have implemented the queries using that data structure.

\subsection{NodeList data structure}

\begin{table}
\centering
\caption{Operations of \textit{NodeList} Data Structure}
\begin{tabular}{|l|l|} \hline
Operation Signature & Time Complexity \\ \hline \hline
add(key : Object, value : NodeValue) & O(log(n)) \\ \hline 
get(key : Object) : NodeValue & O(1) \\ \hline
remove(key : Object) : NodeValue & O(log(n)) \\ \hline
decrementPosition(key : Object) & O(log(n)) \\ \hline
incrementPosition(key : Object) & O(log(n)) \\ \hline
getTopValues() : List<NodeValue> & O(1) \\ \hline
\end{tabular}
\end{table}

### 2015-03-16 paper/query_implementation.tex
adding order list data structure

@@
 \end{tabular}
+\label{nodelist_api}
 \end{table}
+
+Table \ref{nodelist_api} shows the operations of this data structure.
+The NodeValue is an interface to plug any user defined type as a value.
+For an example, the route query can use a RouteCount object to keep route count while profitable cells query can use CellProfict object to keep fare details.
+This interface contains a method called compare to define the order of the values.
+The NodeList data structure provides the notion of a sorted values starting from position 0 (highest value) to its users.
+The decrementPostion operation moves the corresponding value towards position 0 until the correct position according to sorted order.
+Therefore this method should be called after increasing a value of an existing object.
+Similarly incrementPosition operation can be used to move a value to correct position after decreasing the value of an existing object.
+GetTopValues operation returns the top ten values of the list.
+Query algorithm sections further elaborate usage of these methods.
+
+One way of implementing this data structure is to use a heap.
+Heap supports adding and extracting maximum value operations in O(log(n)) time.
+However since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back, if we just a heap.
+On the other hand if we use a doubly linked list to keep values, it will take O(n) time for  decrementPostion and   incrementPosition operations.
+Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and heap to keep other values to avoid above problems.
+A Map which points to values can be used to retrieve values using a key.
+
+NodeList data structure uses two data structures called OrderedList to keep first 10 values and DynamicHeap to keep other values.
+In this way it has to change only the last value of OrderedList and maximum value of Dynamic heap as required.
+Following sections describes how these structures implemented with their operations.
+
+\subsection{OrderedList data structure}
+
+\begin{table}
+\centering
+\caption{Operations of \textit{OrderedList} Data Structure}
+\begin{tabular}{|l|l|} \hline
+Operation Signature & Time Complexity \\ \hline \hline
+add(key : Object, value : NodeValue) & O(n) \\ \hline
+containsKey(key : Object) : boolean & O(1) \\ \hline
+get(key : Object) : NodeValue & O(1) \\ \hline
+incrementPosition(key : Object) & O(n) \\ \hline
+decrementPosition(key : Object) & O(n) \\ \hline
+remove(key : Object) : NodeValue & O(1) \\ \hline
+getLastPosition() : int & O(1) \\ \hline
+getLastKey() : Object & O(1) \\ \hline
+\end{tabular}
+\label{orderedlist_api}
+\end{table}
+
+Table \ref{orderedlist_api} shows the operations of the \textit{OrderedList} data structure.
+First six methods has the same meanings as the \textit{NodeList} data structure and last two can be used to get the size of the list and the value of the last object.
+We internally keep a counter and a pointer to tail to support these operations in O(1) time.
+Figure \ref{orderedlist_impl} shows the implementation details.
+The Map structure (an HashMap) is used to retrieve object pointers with O(1) expected time.
+Each node of the doubly linked list points to previous and next nodes while storing the \textit{NodeValue}.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{orderedlist.png}
+        \caption{\textit{OrderedList} data structure implementation}
+        \label{orderedlist_impl}
+\end{figure}

### 2015-03-16 paper/query_implementation.tex
adding algorithum details

@@
 \end{figure}
+
+\subsection{DynamicHeap data structure}
+
+\begin{table}
+\centering
+\caption{Operations of \textit{DynamicHeap} Data Structure}
+\begin{tabular}{|l|l|} \hline
+Operation Signature & Time Complexity \\ \hline \hline
+add(key : Object, value : NodeValue) & O(log(n)) \\ \hline
+remove(key : Object) : NodeValue & O(log(n)) \\ \hline
+containsKey(key : Object) : boolean & O(1) \\ \hline
+get(key : Object) : NodeValue & O(1) \\ \hline
+moveUp(key : Object) & O(log(n)) \\ \hline
+moveDown(key : Object) & O(log(n)) \\ \hline
+getMaxkey() : Object & O(1) \\ \hline
+extractMax() : NodeValue & O(log(n)) \\ \hline
+\end{tabular}
+\label{dynamicheap_api}
+\end{table}
+
+Table \ref{dynamicheap_api} shows the operations of the \textit{DynamicHeap} data structure.
+\texttt{MoveUp} and \texttt{moveDown} operations are used to move an element up in the heap if the value increased, or move down in the heap if the value decreased.
+Figure \ref{dynamicheap_impl} shows the implementation details.
+Conventional heaps only support \texttt{add} and \texttt{extractMax} operations since heap operations requires to know the element position in heap.
+When adding an element we add it to last position and move that up and extracting maximum element start at array position 0.
+In order to support \texttt{moveUp}, \texttt{moveDown} and \texttt{remove} operations, we keep the array index with the object itself.
+This allows us to retrieve array index in O(1) time and move the element O(log(n)) time.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{DinamicHeap.png}
+        \caption{\textit{DynamicHeap} data structure implementation}
+        \label{dynamicheap_impl}
+\end{figure}
+
+\subsection{Frequent Route Query}
+
+Frequent route query requires to find the top frequent routes within last 30 minutes.
+Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.
+Since queue operations are supported in O(1) time, query evaluation happens in O(log(n)) time.
+Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to process an event.
+This algorithm assumes there is a global NodeList object and route counts stored in an Object called RouteCount.
+
+\begin{algorithm}
+\caption{Algorithm to generate top 10 frequent route change events}
+\label{frequent_route_algorithm}
+\begin{algorithmic} 
+\STATE beforeTopTen $\leftarrow$ nodeList.getTopValues() 
+\STATE window.add(event) 
+\IF{ nodeList.containsKey(event.route) }
+	\STATE routCount $\leftarrow$ nodeList.get(event.route) 
+	\STATE routCount.count++ 
+	\STATE nodeList.decrementPosition(event.route) 
+\ELSE
+	\STATE nodeList.add(event.route, new RouteCount(1)) 
+\ENDIF
+
+\WHILE{ there exists expired events }
+	\STATE  expEvent $\leftarrow$ window.poll() 
+	\STATE  routCount $\leftarrow$ nodeList.get(expEvent.route)
+	\STATE  routCount.count-- 
+	\IF{ routCount.count is 0 }
+		\STATE nodelist.remove(expEvent.route)
+	\ELSE
+		\STATE nodeList.incrementPosition(expEvent.route) 
+	\ENDIF
+\ENDWHILE
+
+\STATE nowTopTen $\leftarrow$ nodeList.getTopValues()
+
+\IF{ preTopTen not equals nowTopTen }
+	\STATE create new event with top ten route details 
+\ENDIF
+
+\end{algorithmic}
+\end{algorithm}
+
+Processing an event can cause add, remove or update values of the existing time window objects.
+Therefore query processing mainly involves with performing above operations to the underlying \textit{NodeList} data structure and compare before and after top ten routes.
+As shown in Algorithm \ref{frequent_route_algorithm}, when an event arrives, first it put the event to time window to keep track for last 30 minutes.
+Then it adds new route value or update the existing value and move its position.
+Similarly, for expired events either it removes the values or reduces the count for existing values.
+Finally, two lists are compared to decide whether to generate an event or not.
+Here we have given a rough out line of the algorithm.
+Real implementation considers other factors such as sequence number to order events.

### 2015-03-16 paper/query_implementation.tex
finishing query evaluation

@@
 Real implementation considers other factors such as sequence number to order events.
+
+\subsection{Profitable Cells Query}
+
+We have implemented profitable cells query in a similar manner to the frequent route query using  time windows. Complexity of the query raised following issues and we have solved them as given bellow.
+\begin{enumerate}
+	\item The profitability is calculated using the median profit and the number of empty taxis.
+In order to calculate median with O(log(n)) time complexity, we use two heaps (java.util.PriorityQueue) .
+One is used to keep the values lesser than the median value (max heap) and other is to keep higher values (min heap).
+However the delete operation of java.util.PriorityQueue takes O(n) time, since it traverse along the array to find the index of the object.
+We tried with our DynamicHeap implementation to fix this.
+But that did not improved the performance.
+Firstly, these heap sizes does not grow over the value of 60 and there are same values repeating in the list increasing the probability of find a value without traversing the whole list.
+Secondly, our implementation create more objects compared to just keeping double values though the time complexity is O(log(n)).
+	\item In order to keep tract of empty taxis, we need to process both pickup cell and drop off cell details for each message.
+One way of doing this is to send all events for same process instance and process both details with one event.
+However this avoid the possibility of partitioning data across different processes.
+We generate two events one for pickup data and one for drop off data to make the process parallel.
+	\item In order to handle expiration of drop off events, we keep track of the taxi ids currently available in a particular cell.
+If a pickup event occurs before drop off event expiration, then we remove that taxi id from that cell and reduce the empty taxi count.
+In such a case drop off event expiration does not cause a reduction of taxi count.
+\end{enumerate}

### 2015-03-16 paper/query_evaluation.tex
adding query evaluation

\section{Query evaluation}

In this section, we explain how to evaluate queries with the given data set. For this query evaluation we can use a distributed event stream processing system with the process graph as shown in Figure \ref{twonode_graph} for both queries.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{twonode_graph.png}
        \caption{Two node process graph to evaluate queries}
        \label{twonode_graph}
\end{figure}

\textit{EventEmitter} reads the data file and generates events. For real implementation, each query has its own \textit{EventEmitter} which reads each line of the data file and generates an event with relevant information. \textit{QueryEvaluator} evaluates the query against the received event and generates top ten value changed event.  In general we can evaluate queries sequentially and parallel. In a sequential execution, all events goes to one \textit{QueryEvaluator} instance and in parallel execution, event stream is partitioned using the key field and evaluate simultaneously with different \textit{QueryEvaluator} instances.

\subsection{Sequential evaluation}

We can sequentially execute events either in a single Java Virtual Machine (JVM) or different JVMs (either in one machine or different machines) using TCP sockets to communicate between process instances. Figure \ref{sequential} shows such possible executions.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{sequential.png}
        \caption{Different types of sequential execution}
        \label{sequential}
\end{figure}

First both processors can be executed in same JVM with same thread. In that case both processes executes sequentially. We can make two processes parallel by executing them with two threads and communicate using a queue in same JVM or communicate using a TCP connection for different JVMs. In all these cases \textit{QueryEvaluator} processes messages sequentially and becomes the bottleneck. Next we show how to parallel \textit{QueryEvaluator} process.

### 2015-03-17 paper/paper.tex
adding query evaluations

@@
 \documentclass{acm_proc_article-sp}
+
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
@@
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
 %A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
 
-\terms{Theory}
-
-\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings
+\keywords{DEBS Grand Challenge, Distributed Event Processing, Time Complexity} % NOT required for Proceedings
 
 \section{Introduction}
-The \textit{proceedings} are the records of a conference.
-ACM seeks to give these conference by-products a uniform,
-high-quality appearance.  To do this, ACM has some rigid
-requirements for the format of the proceedings documents: there
-is a specified format (balanced  double columns), a specified
-set of fonts (Arial or Helvetica and Times Roman) in
-certain specified sizes (for instance, 9 point for body copy),
-a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
-the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
-and (1.9 cm [.75"]) left and right; specified column width
-(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).
+introduction
 
-The good news is, with only a handful of manual
-settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
-and {\texttt{\char'134 alignauthor}} commands, you have
-already used; another, {\texttt{\char'134 balancecolumns}}, will
-be used in your very last run of \LaTeX\ to ensure
-balanced column heights on the last page.}, the \LaTeX\ document
-class file handles all of this for you.
-
-The remainder of this document is concerned with showing, in
-the context of an ``actual'' document, the \LaTeX\ commands
-specifically available for denoting the structure of a
-proceedings paper, rather than with giving rigorous descriptions
-or explanations of such commands.
-
-\section{The {\secit Body} of The Paper}
-Typically, the body of a paper is organized
-into a hierarchical structure, with numbered or unnumbered
-headings for sections, subsections, sub-subsections, and even
-smaller sections.  The command \texttt{{\char'134}section} that
-precedes this paragraph is part of such a
-hierarchy.\footnote{This is the second footnote.  It
-starts a series of three footnotes that add nothing
-informational, but just give an idea of how footnotes work
-and look. It is a wordy one, just so you see
-how a longish one plays out.} \LaTeX\ handles the numbering
-and placement of these headings for you, when you use
-the appropriate heading commands around the titles
-of the headings.  If you want a sub-subsection or
-smaller part to be unnumbered in your output, simply append an
-asterisk to the command name.  Examples of both
-numbered and unnumbered headings will appear throughout the
-balance of this sample document.
-
-Because the entire article is contained in
-the \textbf{document} environment, you can indicate the
-start of a new paragraph with a blank line in your
-input file; that is why this sentence forms a separate paragraph.
-
-\subsection{Type Changes and {\subsecit Special} Characters}
-We have already seen several typeface changes in this sample.  You
-can indicate italicized words or phrases in your text with
-the command \texttt{{\char'134}textit}; emboldening with the
-command \texttt{{\char'134}textbf}
-and typewriter-style (for instance, for computer code) with
-\texttt{{\char'134}texttt}.  But remember, you do not
-have to indicate typestyle changes when such changes are
-part of the \textit{structural} elements of your
-article; for instance, the heading of this subsection will
-be in a sans serif\footnote{A third footnote, here.
-Let's make this a rather short one to
-see how it looks.} typeface, but that is handled by the
-document class file. Take care with the use
-of\footnote{A fourth, and last, footnote.}
-the curly braces in typeface changes; they mark
-the beginning and end of
-the text that is to be in the different typeface.
-
-You can use whatever symbols, accented characters, or
-non-English characters you need anywhere in your document;
-you can find a complete list of what is
-available in the \textit{\LaTeX\
-User's Guide}\cite{Lamport:LaTeX}.
-
-\subsection{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of
-the three are discussed in the next sections.
-
-\subsubsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an
-inline or in-text formula.  It is produced by the
-\textbf{math} environment, which can be
-invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
-construction or with the short form \texttt{\$. . .\$}. You
-can use any of the symbols and structures,
-from $\alpha$ to $\omega$, available in
-\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
-few examples of in-text equations in context. Notice how
-this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsubsection{Display Equations}
-A numbered display equation -- one set off by vertical space
-from the text and centered horizontally -- is produced
-by the \textbf{equation} environment. An unnumbered display
-equation is produced by the \textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols
-and structures available in \LaTeX; this section will just
-give a couple of examples of display equations in context.
-First, consider the equation, shown as an inline equation above:
-\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\subsection{Citations}
-Citations to articles \cite{bowman:reasoning, clark:pct, braams:babel, herlihy:methodology},
-conference
-proceedings \cite{clark:pct} or books \cite{salas:calculus, Lamport:LaTeX} listed
-in the Bibliography section of your
-article will occur throughout the text of your article.
-You should use BibTeX to automatically produce this bibliography;
-you simply need to insert one of several citation commands with
-a key of the item cited in the proper location in
-the \texttt{.tex} file \cite{Lamport:LaTeX}.
-The key is a short reference you invent to uniquely
-identify each work; in this sample document, the key is
-the first author's surname and a
-word from the title.  This identifying key is included
-with each item in the \texttt{.bib} file for your article.
-
-The details of the construction of the \texttt{.bib} file
-are beyond the scope of this sample document, but more
-information can be found in the \textit{Author's Guide},
-and exhaustive details in the \textit{\LaTeX\ User's
-Guide}\cite{Lamport:LaTeX}.
-
-This article shows only the plainest form
-of the citation command, using \texttt{{\char'134}cite}.
-This is what is stipulated in the SIGS style specifications.
-No other citation format is endorsed.
-
-\subsection{Tables}
-Because tables cannot be split across pages, the best
-placement for them is typically the top of the page
-nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and
-the table caption.  The contents of the table itself must go
-in the \textbf{tabular} environment, to
-be aligned properly in rows and columns, with the desired
-horizontal and vertical rules.  Again, detailed instructions
-on \textbf{tabular} material
-is found in the \textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table 1 is included in the input file; compare the
-placement of the table here with the table in the printed
-dvi output of this document.
-
-\begin{table}
-\centering
-\caption{Frequency of Special Characters}
-\begin{tabular}{|c|c|l|} \hline
-Non-English or Math&Frequency&Comments\\ \hline
-\O & 1 in 1,000& For Swedish names\\ \hline
-$\pi$ & 1 in 5& Common in math\\ \hline
-\$ & 4 in 5 & Used in business\\ \hline
-$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-\hline\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of
-the page's live area, use the environment
-\textbf{table*} to enclose the table's contents and
-the table caption.  As with a single-column table, this wide
-table will ``float" to a location deemed more desirable.
-Immediately following this sentence is the point at which
-Table 2 is included in the input file; again, it is
-instructive to compare the placement of the
-table here with the table in the printed dvi
-output of this document.
-
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
-
-\subsection{Figures}
-Like tables, figures cannot be split across pages; the
-best placement for them
-is typically the top or the bottom of the page nearest
-their initial cite.  To ensure this proper ``floating'' placement
-of figures, use the environment
-\textbf{figure} to enclose the figure and its caption.
-
-This sample document contains examples of \textbf{.eps}
-and \textbf{.ps} files to be displayable with \LaTeX.  More
-details on each of these is found in the \textit{Author's Guide}.
-
-\begin{figure}
-\centering
-\epsfig{file=fly.eps}
-\caption{A sample black and white graphic (.eps format).}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=fly.eps, height=1in, width=1in}
-\caption{A sample black and white graphic (.eps format)
-that has been resized with the \texttt{epsfig} command.}
-\end{figure}
-
-As was the case with tables, you may want a figure
-that spans two columns.  To do this, and still to
-ensure proper ``floating'' placement of tables, use the environment
-\textbf{figure*} to enclose the figure and its caption.
-
-Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
-used; use
-the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
-commands as appropriate for the different file types.
-
-\subsection{Theorem-like Constructs}
-Other common constructs that may occur in your article are
-the forms for logical constructs like theorems, axioms,
-corollaries and proofs.  There are
-two forms, one produced by the
-command \texttt{{\char'134}newtheorem} and the
-other by the command \texttt{{\char'134}newdef}; perhaps
-the clearest and easiest way to distinguish them is
-to compare the two in the output of this sample document:
-
-This uses the \textbf{theorem} environment, created by
-the\linebreak\texttt{{\char'134}newtheorem} command:
-\newtheorem{theorem}{Theorem}
-\begin{theorem}
-Let $f$ be continuous on $[a,b]$.  If $G$ is
-an antiderivative for $f$ on $[a,b]$, then
-\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
-\end{theorem}
-
-The other uses the \textbf{definition} environment, created
-by the \texttt{{\char'134}newdef} command:
-\newdef{definition}{Definition}
-\begin{definition}
-If $z$ is irrational, then by $e^z$ we mean the
-unique number which has
-logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
-\end{definition}
-
-Two lists of constructs that use one of these
-forms is given in the
-\textit{Author's  Guidelines}.
-
-\begin{figure*}
-\centering
-\epsfig{file=flies.eps}
-\caption{A sample black and white graphic (.eps format)
-that needs to span two columns of text.}
-\end{figure*}
-and don't forget to end the environment with
-{figure*}, not {figure}!
-
-There is one other similar construct environment, which is
-already set up
-for you; i.e. you must \textit{not} use
-a \texttt{{\char'134}newdef} command to
-create it: the \textbf{proof} environment.  Here
-is a example of its use:
-\begin{proof}
-Suppose on the contrary there exists a real number $L$ such that
-\begin{displaymath}
-\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
-\end{displaymath}
-Then
-\begin{displaymath}
-l=\lim_{x\rightarrow c} f(x)
-= \lim_{x\rightarrow c}
-\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
-= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
-\frac{f(x)}{g(x)} = 0\cdot L = 0,
-\end{displaymath}
-which contradicts our assumption that $l\neq 0$.
-\end{proof}
-
-Complete rules about using these environments and using the
-two different creation commands are in the
-\textit{Author's Guide}; please consult it for more
-detailed instructions.  If you need to use another construct,
-not listed therein, which you want to have the same
-formatting as the Theorem
-or the Definition\cite{salas:calculus} shown above,
-use the \texttt{{\char'134}newtheorem} or the
-\texttt{{\char'134}newdef} command,
-respectively, to create it.
-
-\subsection*{A {\secit Caveat} for the \TeX\ Expert}
-Because you have just been given permission to
-use the \texttt{{\char'134}newdef} command to create a
-new form, you might think you can
-use \TeX's \texttt{{\char'134}def} to create a
-new command: \textit{Please refrain from doing this!}
-Remember that your \LaTeX\ source code is primarily intended
-to create camera-ready copy, but may be converted
-to other forms -- e.g. HTML. If you inadvertently omit
-some or all of the \texttt{{\char'134}def}s recompilation will
-be, to say the least, problematic.
-
-\section{Conclusions}
-This paragraph will end the body of this sample document.
-Remember that you might still have Acknowledgments or
-Appendices; brief samples of these
-follow.  There is still the Bibliography to deal with; and
-we will make a disclaimer about that here: with the exception
-of the reference to the \LaTeX\ book, the citations in
-this paper are to articles which have nothing to
-do with the present subject and are used as
-examples only.
-%\end{document}  % This is where a 'short' article might terminate
-
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.  In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
+\input{query_implementation}
+\input{query_evaluation}
 
@@
 %\balancecolumns
-\appendix
 %Appendix A
-\section{Headings in Appendices}
-The rules about hierarchical headings discussed above for
-the body of the article are different in the appendices.
-In the \textbf{appendix} environment, the command
-\textbf{section} is used to
-indicate the start of each Appendix, with alphabetic order
-designation (i.e. the first is A, the second B, etc.) and
-a title (if you include one).  So, if you need
-hierarchical structure
-\textit{within} an Appendix, start with \textbf{subsection} as the
-highest level. Here is an outline of the body of this
-document in Appendix-appropriate form:
-\subsection{Introduction}
-\subsection{The Body of the Paper}
-\subsubsection{Type Changes and  Special Characters}
-\subsubsection{Math Equations}
-\paragraph{Inline (In-text) Equations}
-\paragraph{Display Equations}
-\subsubsection{Citations}
-\subsubsection{Tables}
-\subsubsection{Figures}
-\subsubsection{Theorem-like Constructs}
-\subsubsection*{A Caveat for the \TeX\ Expert}
-\subsection{Conclusions}
-\subsection{Acknowledgments}
-\subsection{Additional Authors}
-This section is inserted by \LaTeX; you do not insert it.
-You just add the names and information in the
-\texttt{{\char'134}additionalauthors} command at the start
-of the document.
-\subsection{References}
-Generated by bibtex from your ~.bib file.  Run latex,
-then bibtex, then latex twice (to resolve references)
-to create the ~.bbl file.  Insert that ~.bbl file into
-the .tex source file and comment out
-the command \texttt{{\char'134}thebibliography}.
-% This next section command marks the start of
-% Appendix B, and does not continue the present hierarchy
-\section{More Help for the Hardy}
-The acm\_proc\_article-sp document class file itself is chock-full of succinct
-and helpful comments.  If you consider yourself a moderately
-experienced to expert user of \LaTeX, you may find reading
-it useful but please remember not to change it.
-\balancecolumns
 % That's all folks!

### 2015-03-17 paper/query_evaluation.tex
adding query evaluations

@@
         \centering
-        \includegraphics[width=3.0in]{twonode_graph.png}
+        \includegraphics[width=2.0in]{twonode_graph.png}
         \caption{Two node process graph to evaluate queries}
@@
 Next we show how to parallel \textit{QueryEvaluator} process.
+
+\subsection{Parallel evaluation}
+The general way to parallelize event processing, is to partition the events according to key and run multiple instances to process data.
+Therefore in this section, we explain how to parallelize the query evaluation process taking frequent rote query into consideration.
+Same concepts can be applied to profitable cell evaluation as well.
+
+For frequent route query, route events can be partitioned using the hash code of the route object and process different route sets in different \textit{QueryEvaluators}.
+In such a setup, each \textit{QueryEvaluator} generates an event when top ten routes get changed for its route set.
+Since this is not the desired result, we can add another process to aggregate these top result events and generate global route change events.
+Figure \ref{parallel} shows the possible execution patterns with the new process.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{parallel.png}
+        \caption{Different types of parallel execution}
+        \label{parallel}
+\end{figure}
+
+\textit{ResultAggregator} keeps a set of route counts received by \textit{QueryEvaluators}.
+Upon receiving a new event, it can update the existing route count values and check whether global top ten route counts has been changed.
+If there is a change, it can generate a top ten route changed event.
+This logic has the following issues and can be fixed as given bellow.
+
+\subsubsection{Expired values}
+Route count of a particular route can be reduced due to expiration of some events.
+For an example, lets assume route count for route R1 is 10 and R1 is included in a top ten route event generated by a \textit{QueryEvaluator}.
+Then lets assume route count for R1 has been reduced to 7 (due to expiration events), when the next event is generated and hence R1 is not included in the next top ten route event.
+However since R1 value still is in the \textit{ResultAggregator} that may produce an event with this value.
+To avoid this problem, we send the removed route set details with the top ten route event generated by the \textit{QueryEvaluator}.
+In that case \textit{ResultAggregator} can first remove those routes.
+\subsubsection{Out of order events}
+\textit{ResultAggregator} may receive events in a different order than those were produced at the \textit{EventEmitter}.
+This out of order events may change the final output of the \textit{ResultAggregator}.
+We do message ordering at the \textit{ResultAggregator} to avoid this problem.
+In order to order messages, first we kept a message queue for each \textit{QueryEvaluator} and sent the message with least sequence number if a message is present in all queues.
+However this resulted long queues due to speed difference of the different \textit{QueryEvaluator} processes.
+So we limited the queue size for 100.
+If a queue becomes 100, then we send all the messages less than the last message of that queue to \textit{ResultAggregator}.
+\subsubsection{Intermediate top ten route events}
+A route event increases the route count for that events' route.
+Further that can generate some expiration events which may reduce the route count for other routes.
+The top ten frequent routes can be changed due to these expiration events as well.
+Lets say in one \textit{QuerayEvaluator} case, a particular event cause two routes R1 and R2 to change and generates a top ten route change event.
+In the distributed setup these two routes can be in two \textit{QueryEvaluators} and that can cause two top ten route change events.
+These two events may result in two top ten route change events at \textit{ResultAggregator} as well instead of one.
+We could not fixed this issue, but we observe around 5\% of such events generated and system goes to the one \textit{QueryEveluator} result after all such event execution.
+
+Algorithm \ref{aggregate_algorithm} shows the algorithm for ResultAggregator with above fixes.
+First it removes the expired keys.
+Then it update the existing values or add new values and move the position of the list accordingly.
+Finally it generates an event if there is a change to the top list.
+
+\begin{algorithm}
+\caption{Algorithm to aggregate top ten results from QueryEvaluators and generate global events}
+\label{aggregate_algorithm}
+\begin{algorithmic}
+
+\STATE preTopTen $\leftarrow$ nodeList.getTopValues()
+\FORALL {key in event.removedKeys}
+	\STATE nodeList.remove(key)
+\ENDFOR
+
+\FORALL {nodeValue in event.newValuesList}
+	\IF {nodeList.containsKey(nodeValue.key)}
+		\STATE existingValue $\leftarrow$ nodeList.get(nodeValue.key)
+		\IF { existing value is lesser than current }
+			\STATE update the existing value
+			\STATE nodeList.decrementPosition(nodeValue.key)
+		\ELSE
+			\STATE update the existing value
+			\STATE nodeList.incrementPosition(nodeValue.key)
+		\ENDIF
+	\ELSE
+		\STATE nodeList.add(nodeValue.key, nodeValue)
+	\ENDIF
+\ENDFOR
+
+\STATE nowTopTen $\leftarrow$ nodeList.getTopValues()
+\IF { preTopTen not equals nowTopTen }
+	\STATE emit new event
+\ENDIF
+\end{algorithmic}
+\end{algorithm}

### 2015-03-17 paper/experiments.tex
adding experiments

\section{Experiments}
We conduced several experiments to measure the soundness of our solution. Since the parallel message processing gives the better performance, we present details of those experiments with distributed setup as well a singel JVM setup. All these experiments were performed on a LAN with a network bandwidth of 1Gbps and all machines were Intel(R) Xeon(R) 2.4GHz, 4 core duo machines with 16 GB of memory.

\subsection{Frequent routes query}

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{routegraph.png}
        \caption{Process graph to evaluate top ten frequent routes.}
        \label{routegraph}
\end{figure}

Figure \ref{routegraph} shows the process graph used to process frequent route queries. \textit{RouteEventEmitter} reads the data file and emits events with a sequence number. \textit{RouteEmitter} contains a set of threads (equals to the number of \textit{RouteProcessors}) to emit events more efficiently. When processing each line first it reads the data line and creates a Route object using longitude and latitude coordinates of pickup and dropoff locations. Then it distributes the events among the \textit{RouteProcessors} using the hash value of the Route object. \textit{TopRouteProcessor} aggregates the results and saves them to a file.

We deployed this process graph to a distributed stream processing system we have developed, and  measured the execution time and the average message latency varying the number of route processors with the data set given (173 million records). In local set up, we use a new thread to execute a new instance and in distributed setup we add a new node to execute the new instance.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{time_route.png}
        \caption{Execution times to evaluate top ten frequent routes.}
        \label{time_route}
\end{figure}

Figure \ref{time_route} shows the execution times. For both setups execution time is higher for single instance due to lack of parallelism of the system. For distributed setup time get reduced with addition of new nodes and comes to a constant level. For single JVM set up highest throughput is achieved with two route processors. After that point time get slightly increases due to context switching.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{latency_route.png}
        \caption{Latencies to evaluate top ten frequent routes.}
        \label{latency_route}
\end{figure}

Figure \ref{latency_route} shows the latencies. For single JVM setup latency get reduced at 2 and slightly get increase due to context switching as in the execution times. For distributed set up initial latency is high due to communication and message serialization parsing overheads. With higher number of processors latency get reduced.

\subsection{Profitable cells query}

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{profictgraph.png}
        \caption{Process graph to evaluate top ten profitable cells.}
        \label{profictgraph}
\end{figure}

Figure \ref{profictgraph} shows the process graph for profitable cells query. \textit{ProfitEventEmitter} reads the data file and calculates the pickup cell and dropoff cell details using longitude and latitude coordinates. Then it creates two events one for dropoff part and other to pickup part, with different sequence numbers and distributes the events among \textit{ProfictCalculators} using the hash value of the cell. In this way cells are partitioned into different \textit{ProfitCalculator} instances and each instance gets all the events relevant to its' cell set. We use a set of threads (equals to the number of Profict Calculators) to emit data as in the previous query. Finally \textit{TopProfictProcessor} aggregates events and writes to a file. 
Similar to earlier experiment,  we deployed this process graph to our system and measured the execution time and the average message latency with the data set given for both single JVM setup and distributed setup.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{time_profit.png}
        \caption{Execution times to evaluate top ten profitable cells.}
        \label{time_profit}
\end{figure}

Figure \ref{time_profit} shows the execution times for both setups. In both setups, there is a higher latency at one instance due to lack of parallelism. Then for distributed set up execution times get reduced with addition of new resources and for single JVM setup it remains a constant.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{latency_profit.png}
        \caption{Latencies to evaluate top ten profitable cells.}
        \label{latency_profit}
\end{figure}

Figure \ref{latency_profit} shows the latencies. For single JVM, latencies slightly increased with the addition of new instances as expected. But for distributed setups there is a higher latencies compared to others. This latency is due to longer queues at the \textit{TopProfitProcessor}.

### 2015-03-17 paper/paper.tex
adding experiments

@@
 Then we evaluate our solution within signal node as well as in distributed setups.
-We could process frequent routes query with 173 million data set within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
+We could process frequent routes query with 173 million dataset within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
 \end{abstract}
@@
 \input{query_evaluation}
+\input{experiments}
 

### 2015-03-17 paper/query_evaluation.tex
adding experiments

@@
 
-In this section, we explain how to evaluate queries with the given data set.
+In this section, we explain how to evaluate queries with the given dataset.
 For this query evaluation we can use a distributed event stream processing system with the process graph as shown in Figure \ref{twonode_graph} for both queries.
@@
 
-\textit{EventEmitter} reads the data file and generates events.
+\textit{EventEmitter} reads the data file and generates events with a sequence number.
+This sequence number is used to order \textit{QueryEvaluator} results.
 For real implementation, each query has its own \textit{EventEmitter} which reads each line of the data file and generates an event with relevant information.
@@
 First both processors can be executed in same JVM with same thread.
-In that case both processes executes sequentially.
+In that case, both processes executes sequentially.
 We can make two processes parallel by executing them with two threads and communicate using a queue in same JVM or communicate using a TCP connection for different JVMs.

### 2015-03-17 paper/query_implementation.tex
adding experiments

@@
 Further since each event is received with route and cell details, above mentioned operations should perform using route and cell details as the key.
-In order to support all these functions with O(log(n)) time complexity we developed a data structure called NodeList.
+In order to support all these functions with O(log(n)) time complexity we developed a data structure called \textit{NodeList}.
 Following parts of this section describes how we have achieve this time complexity with \textit{NodeList} data structure and how we have implemented the queries using that data structure.
@@
 Table \ref{nodelist_api} shows the operations of this data structure.
-The NodeValue is an interface to plug any user defined type as a value.
-For an example, the route query can use a RouteCount object to keep route count while profitable cells query can use CellProfict object to keep fare details.
+The \textit{NodeValue} is an interface to plug any user defined type as a value.
+For an example, the route query can use a \textit{RouteCount} object to keep route count while profitable cells query can use \textit{CellProfict} object to keep fare details.
 This interface contains a method called compare to define the order of the values.
-The NodeList data structure provides the notion of a sorted values starting from position 0 (highest value) to its users.
-The decrementPostion operation moves the corresponding value towards position 0 until the correct position according to sorted order.
+The \textit{NodeList} data structure provides the notion of a sorted values starting from position 0 (highest value) to its users.
+The \texttt{decrementPostion} operation moves the corresponding value towards position 0 until the correct position according to sorted order.
 Therefore this method should be called after increasing a value of an existing object.
-Similarly incrementPosition operation can be used to move a value to correct position after decreasing the value of an existing object.
-GetTopValues operation returns the top ten values of the list.
+Similarly \texttt{incrementPosition} operation can be used to move a value to correct position after decreasing the value of an existing object.
+\texttt{GetTopValues} operation returns the top ten values of the list.
 Query algorithm sections further elaborate usage of these methods.
@@
 Heap supports adding and extracting maximum value operations in O(log(n)) time.
-However since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back, if we just a heap.
-On the other hand if we use a doubly linked list to keep values, it will take O(n) time for  decrementPostion and   incrementPosition operations.
-Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and heap to keep other values to avoid above problems.
+However since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back.
+On the other hand if we use a doubly linked list to keep values, it will take O(n) time for  \texttt{decrementPostion} and \texttt{incrementPosition} operations.
+Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.
 A Map which points to values can be used to retrieve values using a key.
 
-NodeList data structure uses two data structures called OrderedList to keep first 10 values and DynamicHeap to keep other values.
-In this way it has to change only the last value of OrderedList and maximum value of Dynamic heap as required.
-Following sections describes how these structures implemented with their operations.
+\textit{NodeList} data structure uses two data structures called \textit{OrderedList} to keep first 10 values and \textit{DynamicHeap} to keep other values.
+In this way it has to change only the last value of \textit{OrderedList} and maximum value of \textit{DynamicHeap} as required.
+Following sections describes how these data structures have been implemented with their operations.
 
@@
 Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to process an event.
-This algorithm assumes there is a global NodeList object and route counts stored in an Object called RouteCount.
+This algorithm assumes there is a global \textit{NodeList} object and route counts stored in an Object called \textit{RouteCount}.
 
@@
 Therefore query processing mainly involves with performing above operations to the underlying \textit{NodeList} data structure and compare before and after top ten routes.
-As shown in Algorithm \ref{frequent_route_algorithm}, when an event arrives, first it put the event to time window to keep track for last 30 minutes.
-Then it adds new route value or update the existing value and move its position.
+As shown in Algorithm \ref{frequent_route_algorithm}, when an event arrives, first it puts the event to time window to keep track for last 30 minutes.
+Then it adds new route value or updates the existing value and moves its position.
 Similarly, for expired events either it removes the values or reduces the count for existing values.
 Finally, two lists are compared to decide whether to generate an event or not.
-Here we have given a rough out line of the algorithm.
+Here we have given a rough outline of the algorithm.
 Real implementation considers other factors such as sequence number to order events.
@@
 	\item The profitability is calculated using the median profit and the number of empty taxis.
-In order to calculate median with O(log(n)) time complexity, we use two heaps (java.util.PriorityQueue) .
+In order to calculate median with O(log(n)) time complexity, we use two heaps (\textit{java.util.PriorityQueue}) .
 One is used to keep the values lesser than the median value (max heap) and other is to keep higher values (min heap).
-However the delete operation of java.util.PriorityQueue takes O(n) time, since it traverse along the array to find the index of the object.
-We tried with our DynamicHeap implementation to fix this.
+However the \texttt{delete} operation of \textit{java.util.PriorityQueue} takes O(n) time, since it traverse along the array to find the index of the object.
+We tried with our \textit{DynamicHeap} implementation to fix this.
 But that did not improved the performance.
@@
 Secondly, our implementation create more objects compared to just keeping double values though the time complexity is O(log(n)).
-	\item In order to keep tract of empty taxis, we need to process both pickup cell and drop off cell details for each message.
+	\item In order to keep tract of empty taxis, we need to process both pickup cell and dropoff cell details for each message.
 One way of doing this is to send all events for same process instance and process both details with one event.
 However this avoid the possibility of partitioning data across different processes.
-We generate two events one for pickup data and one for drop off data to make the process parallel.
-	\item In order to handle expiration of drop off events, we keep track of the taxi ids currently available in a particular cell.
-If a pickup event occurs before drop off event expiration, then we remove that taxi id from that cell and reduce the empty taxi count.
-In such a case drop off event expiration does not cause a reduction of taxi count.
+We generate two events, one for pickup data and one for dropoff data to make the process parallel.
+	\item In order to handle expiration of dropoff events, we keep track of the taxi ids currently available in a particular cell.
+If a pickup event occurs before dropoff event expiration, then we remove that taxi id from that cell and reduce the empty taxi count.
+In such a case dropoff event expiration does not cause a reduction of taxi count.
 \end{enumerate}

### 2015-03-18 paper/paper.tex
adding related work

@@
 
-\title{Solving the grand challenge using an algorithmic an distributed approach}
+\title{Solving the grand challenge using an algorithmic and distributed approach}
 %
@@
 \input{experiments}
-
+\input{related_work}
 %
@@
 \bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
 % You must have a proper ".bib" file

### 2015-03-18 paper/related_work.tex
adding related work

\section{Related work}
DEBS Grand Challenge problem naturally fallen into distributed stream processing domain. Many such systems have been developed with different feature sets according to the applications they use. Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track the trends in web queries. All these system are based on Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data. Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence dynamic membership management. Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets RDDs \cite{zaharia2012resilient} and Muppet \cite{lam2012muppet} provides state persistence using key value store based on Cassandra \cite{lakshman2010cassandra}.

Several approaches has been developed for DEBS Grand Challenge 2014 \cite{jerzak2014debs}. Predictive Load Management in Smart Grid Environments \cite{mutschler2014predictive} provides a solutions based on EventCore. They have developed an efficient algorithms to find median. Scalable Stateful Stream Processing for Samart Grids \cite{fernandez2014scalable} implemented their solution on SEEP \cite{castro2013integrating} a scalable fault tolerant framework. Solving the Grand Challenge Using an Opensource CEP Engine \cite{perera2014solving} solution is based on WSO2 Complex Event Processor. All these solutions partition the data to scale up the application, provide solutions to predict some missing data and provides algorithmic solutions for efficient query processing. Our solution shares the same ideas along these solutions while addressing the specific issues for 2015 Grand Challenge.

### 2015-03-18 src/main/java/edu/colostate/cs/gc/list/DynamicHeap.java
adding a function to find median in log n time

@@ -12,12 +12,29 @@ import java.util.Map;
  */
 public class DynamicHeap {
 
-    private HeapNode[] heapNodes = new HeapNode[20000];
-    private Map<Object, HeapNode> keyMap = new HashMap<Object, HeapNode>();
+    private HeapNode[] heapNodes;
+    private Map<Object, HeapNode> keyMap;
+    private int capacity = 0;
 
     private int heapSize = 0;
 
+    public DynamicHeap() {
+        this.heapNodes = new HeapNode[20000];
+        this.keyMap = new HashMap<Object, HeapNode>(500);
+        this.capacity = 20000;
+
+    }
+
+    public DynamicHeap(int capacity){
+        this.heapNodes = new HeapNode[capacity];
+        this.keyMap = new HashMap<Object, HeapNode>(capacity);
+        this.capacity = capacity;
+    }
+
     public void add(Object key, NodeValue value) {
+        if (this.heapSize == this.capacity){
+           increaseCapacity();
+        }
         HeapNode heapNode = new HeapNode(value, key);
         this.keyMap.put(key, heapNode);
         this.heapNodes[this.heapSize] = heapNode;
@@ -25,6 +42,19 @@ public class DynamicHeap {
         this.heapSize++;
         moveUp(heapNode);
 
+    }
+
+    private void increaseCapacity(){
+        this.capacity = this.capacity * 2;
+        HeapNode[] newHeap = new HeapNode[this.capacity];
+        for (int i = 0; i < this.heapSize;i++){
+            newHeap[i] = this.heapNodes[i];
+        }
+        this.heapNodes = newHeap;
+    }
+
+    public boolean isEmpty(){
+        return this.heapSize == 0;
     }
 
     public NodeValue remove(Object key) {
@@ -45,6 +75,7 @@ public class DynamicHeap {
             this.heapSize--;
         }
 
+        this.heapNodes[this.heapSize] = null;
         return heapNode.getNodeValue();
     }
 
@@ -60,13 +91,29 @@ public class DynamicHeap {
         if (this.heapSize == 0) {
             return null;
         }
+
         HeapNode maxElement = this.heapNodes[0];
-        this.heapNodes[0] = this.heapNodes[this.heapSize - 1];
-        this.heapNodes[0].setArrayPosition(0);
-        this.heapSize--;
-        moveDown(this.heapNodes[0]);
+        if (this.heapSize == 1){
+            // i.e we have only one element
+            this.heapSize--;
+        } else {
[…]

### 2015-03-18 src/main/java/edu/colostate/cs/gc/process/TopEventProcessor.java

@@ -112,21 +112,21 @@ public class TopEventProcessor extends TripProcessor {
         }
 
         // add new values
-        for (NodeValue routeCount : topEvent.getNewValueList()) {
+        for (NodeValue nodeValue : topEvent.getNewValueList()) {
 
-            if (!this.nodeList.containsKey(routeCount.getKey())) {
+            if (!this.nodeList.containsKey(nodeValue.getKey())) {
                 // need to create a new object to avoid conflicts with earlier process objects.
-                this.nodeList.add(routeCount.getKey(), routeCount);
+                this.nodeList.add(nodeValue.getKey(), nodeValue);
             } else {
-                NodeValue existingValue = this.nodeList.get(routeCount.getKey());
-                if (routeCount.compare(existingValue) == -1) {
-                    existingValue.update(routeCount);
+                NodeValue existingValue = this.nodeList.get(nodeValue.getKey());
+                if (nodeValue.compare(existingValue) == -1) {
+                    existingValue.update(nodeValue);
                     //if the new value is less it has to move further down.
-                    this.nodeList.incrementPosition(routeCount.getKey());
+                    this.nodeList.incrementPosition(nodeValue.getKey());
                 } else {
                     // if it is not less either can same or increased.
-                    existingValue.update(routeCount);
-                    this.nodeList.decrementPosition(routeCount.getKey());
+                    existingValue.update(nodeValue);
+                    this.nodeList.decrementPosition(nodeValue.getKey());
                 }
             }
         }

### 2015-03-18 src/main/java/edu/colostate/cs/gc/profit/ProfitCellNode.java

@@ -3,7 +3,10 @@ package edu.colostate.cs.gc.profit;
 import edu.colostate.cs.gc.event.Cell;
 import edu.colostate.cs.gc.list.NodeValue;
 
-import java.util.*;
+import java.util.Comparator;
+import java.util.HashSet;
+import java.util.PriorityQueue;
+import java.util.Set;
 
 /**
  * Created with IntelliJ IDEA.

### 2015-03-18 src/main/java/edu/colostate/cs/gc/profit/ProfitEventEmitter.java

@@ -75,6 +75,7 @@ public class ProfitEventEmitter implements Adaptor {
             System.out.println("Start processing messages ...");
             long currentTime = System.currentTimeMillis();
 
+
             while ((line = bufferedReader.readLine()) != null) {
 
                 try {
@@ -93,7 +94,8 @@ public class ProfitEventEmitter implements Adaptor {
 
                     if ((pickUpEvent.getPickUpCell() != null) && (pickUpEvent.getDropOffCell() != null)) {
 
-                        pickUpEvent.setFare(Double.parseDouble(values[11]) + Double.parseDouble(values[14]));
+                        double fare = Double.parseDouble(values[11]) + Double.parseDouble(values[14]);
+                        pickUpEvent.setFare(Math.round(fare * 100)/ 100.0);
                         pickUpEvent.setPayEvent(true);
 
                         int bufferNumber;
@@ -118,6 +120,8 @@ public class ProfitEventEmitter implements Adaptor {
                             bufferNumber = dropOffEvent.getDropOffCell().hashCode() % messageBuffers.length;
                             messageBuffers[bufferNumber].addMessage(dropOffEvent);
                         }
+
+
                     }
 
                 } catch (Exception e) {

### 2015-03-18 src/main/java/edu/colostate/cs/gc/profit/median/MaxHeapValue.java

@@ -1,0 +1,69 @@
+package edu.colostate.cs.gc.profit.median;
+
+import edu.colostate.cs.gc.list.NodeValue;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/18/15
+ * Time: 9:58 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class MaxHeapValue implements NodeValue {
+
+    private int count;
+    private double value;
+
+    public MaxHeapValue(int count, double value) {
+        this.count = count;
+        this.value = value;
+    }
+
+    public int compare(NodeValue value) {
+        MaxHeapValue newValue = (MaxHeapValue) value;
+        if (this.value < newValue.value){
+            return 1;
+        } else if (this.value > newValue.value){
+            return -1;
+        } else {
+            return 0;
+        }
+
+    }
+
+    public void incrementCount(){
+        this.count++;
+    }
+
+    public void decrementCount(){
+        this.count--;
+    }
+
+    public int getCount() {
+        return count;
+    }
+
+    public void setCount(int count) {
+        this.count = count;
+    }
+
+    public double getValue() {
+        return value;
+    }
+
+    public void setValue(double value) {
+        this.value = value;
+    }
+
+    public Object getKey() {
+        return null;  //To change body of implemented methods use File | Settings | File Templates.
+    }
+
+    public void update(NodeValue value) {
+        //To change body of implemented methods use File | Settings | File Templates.
+    }
+
+    public NodeValue getClone() {
+        return null;  //To change body of implemented methods use File | Settings | File Templates.
+    }
+}

### 2015-03-18 src/main/java/edu/colostate/cs/gc/profit/median/MedianFinder.java

@@ -1,0 +1,203 @@
+package edu.colostate.cs.gc.profit.median;
+
+import edu.colostate.cs.gc.list.DynamicHeap;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/18/15
+ * Time: 9:50 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class MedianFinder {
+
+    private DynamicHeap minHeap = new DynamicHeap(8);
+    private DynamicHeap maxHeap = new DynamicHeap(8);
+
+    private double median;
+    // this value is equals to twice of the median field size
+    private int medianCount;
+
+    // position of the median field
+    private int medianPosition;
+
+    // this means median is in middle of two buckets. so real median will be the middle value.
+    private boolean isInMiddle;
+
+    public MedianFinder() {
+        this.median = 0;
+        this.isInMiddle = true;
+    }
+
+    public double getMedian() {
+        return this.median;
+    }
+
+    public void add(double value) {
+        if (this.isInMiddle) {
+            // in this case we don't care about the equality since the median is a fake value. and there is no
+            // possibility that both are equal since we have unique buckets.
+            if (value > this.median) {
+                addToMaxHeap(value);
+                // new median will be the minimum value of this map
+                pullMedianFromMaxHeap();
+            } else {
+                addToMinHeap(value);
+                pullMedianFromMinHeap();
+            }
+            this.isInMiddle = false;
+        } else {
+            // i.e there is a median class available.
+            // when adding an element we assume it adds to end of the bucket. so need to remove from end as well.
+            if (value == this.median) {
+                this.medianCount = this.medianCount + 2;
+                this.medianPosition++;
+            } else if (value > this.median) {
+                addToMaxHeap(value);
+                // median position should advance by one.
+                this.medianPosition++;
+                if (this.medianPosition > this.medianCount) {
+                    //i.e median should in between.
+                    pushMedianToMinHeap();
+                    this.isInMiddle = true;
+                }
+            } else {
+                // i.e current value is less than median.
+                addToMinHeap(value);
+                this.medianPosition--;
+                if (this.medianPosition < 2) {
+                    // i.e median should be in between
+                    pushMedianToMaxHeap();
+                    this.isInMiddle = true;
[…]

### 2015-03-18 src/main/java/edu/colostate/cs/gc/profit/median/MinHeapValue.java

@@ -1,0 +1,69 @@
+package edu.colostate.cs.gc.profit.median;
+
+import edu.colostate.cs.gc.list.NodeValue;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/18/15
+ * Time: 9:58 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class MinHeapValue implements NodeValue {
+
+    private int count;
+    private double value;
+
+    public MinHeapValue(int count, double value) {
+        this.count = count;
+        this.value = value;
+    }
+
+    public int compare(NodeValue value) {
+        MinHeapValue newValue = (MinHeapValue) value;
+        if (this.value > newValue.value){
+            return 1;
+        } else if (this.value < newValue.value){
+            return -1;
+        } else {
+            return 0;
+        }
+
+    }
+
+    public void incrementCount(){
+        this.count++;
+    }
+
+    public void decrementCount(){
+        this.count--;
+    }
+
+    public int getCount() {
+        return count;
+    }
+
+    public void setCount(int count) {
+        this.count = count;
+    }
+
+    public double getValue() {
+        return value;
+    }
+
+    public void setValue(double value) {
+        this.value = value;
+    }
+
+    public Object getKey() {
+        return this.value;
+    }
+
+    public void update(NodeValue value) {
+        //To change body of implemented methods use File | Settings | File Templates.
+    }
+
+    public NodeValue getClone() {
+        return null;  //To change body of implemented methods use File | Settings | File Templates.
+    }
+}

### 2015-03-18 src/test/java/edu/colostate/cs/gc/TestClass.java

@@ -24,4 +24,6 @@ public class TestClass implements Comparable<TestClass> {
             return 0;
         }
     }
+
+
 }

### 2015-03-18 src/test/java/edu/colostate/cs/gc/profit/median/TestMedianFinder.java

@@ -1,0 +1,196 @@
+package edu.colostate.cs.gc.profit.median;
+
+import junit.framework.TestCase;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/18/15
+ * Time: 11:09 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class TestMedianFinder extends TestCase {
+
+    public void testAddUnique(){
+
+        MedianFinder medianFinder = new MedianFinder();
+        medianFinder.add(20.0);
+        assertEquals(medianFinder.getMedian(), 20.0);
+
+        medianFinder.add(25.0);
+        assertEquals(medianFinder.getMedian(), 22.5);
+
+        medianFinder.add(17.0);
+        assertEquals(medianFinder.getMedian(), 20.0);
+
+        medianFinder.add(14.0);
+        assertEquals(medianFinder.getMedian(), 18.5);
+
+        medianFinder.add(22.0);
+        medianFinder.add(18.0);
+        assertEquals(medianFinder.getMedian(), 19.0);
+    }
+
+    public void testAddDuplicate(){
+        MedianFinder medianFinder = new MedianFinder();
+        medianFinder.add(25.0);
+        assertEquals(medianFinder.getMedian(), 25.0);
+        medianFinder.add(25.0);
+        assertEquals(medianFinder.getMedian(), 25.0);
+        medianFinder.add(26.0);
+        assertEquals(medianFinder.getMedian(), 25.0);
+        medianFinder.add(25.0);
+        assertEquals(medianFinder.getMedian(), 25.0);
+        medianFinder.add(26.0);
+        assertEquals(medianFinder.getMedian(), 25.0);
+        medianFinder.add(26.0);
+        assertEquals(medianFinder.getMedian(), 25.5);
+        medianFinder.add(26.0);
+        assertEquals(medianFinder.getMedian(), 26.0);
+        medianFinder.add(26.0);
+        assertEquals(medianFinder.getMedian(), 26.0);
+        medianFinder.add(24.0);
+        assertEquals(medianFinder.getMedian(), 26.0);
+        medianFinder.add(24.0);
+        assertEquals(medianFinder.getMedian(), 25.5);
+        medianFinder.add(24.0);
+        assertEquals(medianFinder.getMedian(), 25.0);
+
+        medianFinder.add(24.0);
+        assertEquals(medianFinder.getMedian(), 25.0);
[…]

### 2015-03-19 paper/experiments.tex
fixing some issues

@@
 For distributed setup time get reduced with addition of new nodes and comes to a constant level.
+This means our solution scales up to 4 node level and maintains a constant throughput.
+This is due to two reasons.
+Firstly, \textit{RouteProcessor} is not a computing intensive process.
+Secondly, there is only one \textit{RouteEventEmitter} and the speed at which it can read and send messages becomes the bottleneck.
 For single JVM set up highest throughput is achieved with two route processors.
@@
 Then for distributed set up execution times get reduced with addition of new resources and for single JVM setup it remains a constant.
+This shows our solution scales up in a distributed setup.
+We can increase the system throughput by adding new nodes until \textit{ProfitEventEmitter} becomes the bottleneck.
 
@@
 This latency is due to longer queues at the \textit{TopProfitProcessor}.
+In our message ordering process, we can only send a message to \textit{TopProfitProcessor} only if all queues are not empty.
+This allows some processes with less top ten profitable events to delay other process messages.
+This problem can  possibly fixed by adding a limit to the message queue size.
+On the other hand in a real time system, this event ordering many not required since event processing happens at real time.

### 2015-03-19 paper/paper.tex
fixing some issues

@@
 
-\title{Solving the grand challenge using an algorithmic and distributed approach}
+\title{DBES Grand Challenge : High throughput, scalable solution for processing taxi trip data}
 %
@@
 Firstly, such processing demands efficient data structures and algorithms to minimize the processing over head.
-Secondly, system should partition data to process them in parallel to efficient use of hardware.
+Secondly, system should partition data to process them in parallel to make system scalable.
 

### 2015-03-19 paper/query_evaluation.tex
fixing some issues

@@
 
-\subsubsection{Expired values}
-Route count of a particular route can be reduced due to expiration of some events.
+\begin{enumerate}
+	\item \textbf{Expired values} : Route count of a particular route can be reduced due to expiration of some events.
 For an example, lets assume route count for route R1 is 10 and R1 is included in a top ten route event generated by a \textit{QueryEvaluator}.
@@
 In that case \textit{ResultAggregator} can first remove those routes.
-\subsubsection{Out of order events}
-\textit{ResultAggregator} may receive events in a different order than those were produced at the \textit{EventEmitter}.
+	\item \textbf{Out of order events} : \textit{ResultAggregator} may receive events in a different order than those are produced at the \textit{EventEmitter}.
 This out of order events may change the final output of the \textit{ResultAggregator}.
-We do message ordering at the \textit{ResultAggregator} to avoid this problem.
-In order to order messages, first we kept a message queue for each \textit{QueryEvaluator} and sent the message with least sequence number if a message is present in all queues.
-However this resulted long queues due to speed difference of the different \textit{QueryEvaluator} processes.
-So we limited the queue size for 100.
-If a queue becomes 100, then we send all the messages less than the last message of that queue to \textit{ResultAggregator}.
-\subsubsection{Intermediate top ten route events}
-A route event increases the route count for that events' route.
+We order messages at the \textit{ResultAggregator} to avoid this problem.
+In order to order messages, we keep a message queue for each \textit{QueryEvaluator} and sent the message with least sequence number to \textit{ResultAggregator},  if a message is present in all queues.
+	\item \textbf{Intermediate top ten route events} : A route event increases the route count for that events' route.
 Further that can generate some expiration events which may reduce the route count for other routes.
@@
 These two events may result in two top ten route change events at \textit{ResultAggregator} as well instead of one.
-We could not fixed this issue, but we observe around 5\% of such events generated and system goes to the one \textit{QueryEveluator} result after all such event execution.
+Furether we observe only around 5\% of such events generated, and system goes to the one \textit{QueryEveluator} result after all such event are being executed.
+Solution for such a problem may depend on the nature in which a real application, uses these results in decision making process.
+Therefore we could not provide any fix for this problem.
+\end{enumerate}
 

### 2015-03-19 paper/query_implementation.tex
fixing some issues

@@
 But that did not improved the performance.
-Firstly, these heap sizes does not grow over the value of 60 and there are same values repeating in the list increasing the probability of find a value without traversing the whole list.
+Firstly, these heap sizes does not grow over the value of 145 and there are same values repeating in the list increasing the probability of find a value without traversing the whole list.
 Secondly, our implementation create more objects compared to just keeping double values though the time complexity is O(log(n)).

### 2015-03-19 paper/experiments.tex
updated the paper

@@
 Figure \ref{time_profit} shows the execution times for both setups.
-In both setups, there is a higher latency at one instance due to lack of parallelism.
+In both setups, there is a higher execution time at one instance due to lack of parallelism.
 Then for distributed set up execution times get reduced with addition of new resources and for single JVM setup it remains a constant.
@@
 This latency is due to longer queues at the \textit{TopProfitProcessor}.
-In our message ordering process, we can only send a message to \textit{TopProfitProcessor} only if all queues are not empty.
+In our message ordering process, we can send a message to \textit{TopProfitProcessor} only if all queues are not empty.
 This allows some processes with less top ten profitable events to delay other process messages.

### 2015-03-19 paper/query_evaluation.tex
updated the paper

@@
 
-We can sequentially execute events either in a single Java Virtual Machine (JVM) or different JVMs (either in one machine or different machines) using TCP sockets to communicate between process instances.
+We can sequentially execute events either in a single Java Virtual Machine (JVM) or different JVMs (either in one machine or different machines) using TCP sockets to communicate among process instances.
 Figure \ref{sequential} shows such possible executions.
@@
 These two events may result in two top ten route change events at \textit{ResultAggregator} as well instead of one.
-Furether we observe only around 5\% of such events generated, and system goes to the one \textit{QueryEveluator} result after all such event are being executed.
+Further we observed only around 5\% of such events are generated, and system goes to the one \textit{QueryEveluator} result after all such event are being executed.
 Solution for such a problem may depend on the nature in which a real application, uses these results in decision making process.

### 2015-03-19 paper/query_implementation.tex
updated the paper

@@
 Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.
-A Map which points to values can be used to retrieve values using a key.
+A \textit{Map} which points to values can be used to retrieve values using a key.
 
@@
 Figure \ref{orderedlist_impl} shows the implementation details.
-The Map structure (an HashMap) is used to retrieve object pointers with O(1) expected time.
+The \textit{Map} structure (an \textit{java.util.HashMap}) is used to retrieve object pointers with O(1) expected time.
 Each node of the doubly linked list points to previous and next nodes while storing the \textit{NodeValue}.
@@
 Conventional heaps only support \texttt{add} and \texttt{extractMax} operations since heap operations requires to know the element position in heap.
-When adding an element we add it to last position and move that up and extracting maximum element start at array position 0.
+When adding an element, we add it to last position and move that up and extracting maximum element start at array position 0.
 In order to support \texttt{moveUp}, \texttt{moveDown} and \texttt{remove} operations, we keep the array index with the object itself.
@@
 \begin{algorithmic}
+\small 
 \STATE beforeTopTen $\leftarrow$ nodeList.getTopValues() 
@@
 \end{enumerate}
+
+Algorithm \ref{pickup_algorithm} shows the logic to process pickup data.
+When a taxi gets a pickup, the taxi fare is added to that cell.
+We use \textit{ProfitCellNode} to keep the fare and empty taxi details to calculate the profitability.
+Then if its' taxi dropoff event has not been occurred (i.e this taxi has come to this location less than 30 minutes ago) it removes the taxi from that location.
+When the event get expired, it removes the fare from the \textit{ProfitCellNode}.
+
+\begin{algorithm}
+\caption{Algorithm to process pickup data}
+\label{pickup_algorithm}
+\begin{algorithmic}
+\small
+\STATE beforeTopTen $ \leftarrow $ nodeList.getTopValues()
+\STATE paymentWindow.add(event)
+\IF {nodeList.containsKey(event.pickUpCell)}
+	\STATE profitCellNode $ \leftarrow $ nodeList.get(event.pickUpCell)
+	\STATE profitCellNode.addFare(event.fare)
+	\IF {profitCellNode.containsTaxi(event.medalliion)}
+		\STATE profitCellNode.emptyTaxi--
+		\STATE profitCellNode.removeTaxi(event.medallion)
+	\ENDIF
+	\IF {profitablity has increased}
+		\STATE nodeList.decrementPosition(event.pickUpCell)
+	\ELSE
+		\STATE nodeList.incrementPosition(event.pickUpCell)	
+	\ENDIF
+\ELSE
+	\STATE nodeList.add(event.pickUpCell, new ProfitCellNode(event.fare))
+\ENDIF
+\WHILE {there exits expired events}
+	\STATE expiredEvent $ \leftarrow $ paymentWindow.poll()
+	\STATE profitCellNode $ \leftarrow $ nodeList.get(expiredEvent.pickUpCell)
+	\STATE profitCellNode.removeFare(expiredEvent.fare)
+
+	\IF {profitCellNode is empty}
+		\STATE nodeList.remove(expiredEvent.pickUpCell)
+	\ELSIF {profitablity increased}
+		\STATE nodeList.decrementPosition(event.pickUpCell)
+	\ELSE
+		\STATE nodeList.incrementPosition(event.pickUpCell)
+	\ENDIF
+\ENDWHILE
+\STATE nowTopTen $ \leftarrow $ nodeList.getTopValues()
+\IF {beforeTopTen not equals to nowTopTen}
+	\STATE emit a top ten change event
+\ENDIF
+
+\end{algorithmic}
+\end{algorithm}
+
+Algorithm \ref{dropoff_algorithm} shows the logic to process dropoff data.
+When a taxi trip ended in a location, number of empty taxis of that location has to be increased.
+If the taxi does not get a pickup within 30 mintes, then that taxi has to be removed from that location.
+
+\begin{algorithm}
+\caption{Algorithm to process dropoff data}
+\label{dropoff_algorithm}
+\begin{algorithmic}
+\small
+\STATE beforeTopTen $ \leftarrow $ nodeList.getTopValues()
+\STATE dropWindow.add(event)
+\IF {nodeList.containsKey(event.dropOffCell)}
+	\STATE profitCellNode $ \leftarrow $ nodeList.get(event.dropOffCell)
+	\STATE profitCellNode.addTaxi(event.medallion)
+	\STATE profitCellNode.emptyTaxi++
+	\STATE nodeList.incrementPosition(event.dropOffCell)
+
+\ELSE
+	\STATE nodeList.add(event.dropOffCell, new ProfitCellNode(event.medallion))
+\ENDIF
+\WHILE {there are expired events}
+	\STATE expiredEvent $ \leftarrow $ dropWindow.poll()
+	\STATE profitCellNode $ \leftarrow $ nodeList.get(expiredEvent.dropOffCell)
+	\IF {profitCellNode.containsTaxi(expiredEvent.medallion)}
+		\STATE profitCellNode.emptyTaxi--
+		\STATE profitCellNode.removeTaxi(expiredEvent.medallion)
+
+		\IF {profitCellNode is empty}
+			\STATE nodeList.remove(event.dropoffCell)
+		\ELSE 
+			\STATE nodeList.decrementPosition(event.dropOffCell) 
+		\ENDIF
+	\ENDIF
+\ENDWHILE
+
+\end{algorithmic}
+\end{algorithm}

### 2015-03-20 src/main/java/edu/colostate/cs/gc/event/PaymentEvent.java
adding some fixes

@@ -49,6 +49,7 @@ public class PaymentEvent extends TripEvent {
     @Override
     public void serialize(DataOutput dataOutput) throws MessageProcessingException {
         try {
+            dataOutput.writeInt(this.seqNo);
             dataOutput.writeUTF(this.medallion);
             dataOutput.writeDouble(this.fare);
             dataOutput.writeBoolean(this.isPayEvent);
@@ -66,6 +67,7 @@ public class PaymentEvent extends TripEvent {
     @Override
     public void parse(DataInput dataInput) throws MessageProcessingException {
         try {
+            this.seqNo = dataInput.readInt();
             this.medallion = dataInput.readUTF();
             this.fare = dataInput.readDouble();
             this.isPayEvent = dataInput.readBoolean();

### 2015-03-20 src/main/java/edu/colostate/cs/gc/list/NodeList.java

@@ -13,9 +13,8 @@ public class NodeList {
     private OrderedList list;
     private DynamicHeap heap;
 
-
-    public NodeList() {
-        this.heap = new DynamicHeap();
+    public NodeList(int capacity) {
+        this.heap = new DynamicHeap(capacity);
         this.list = new OrderedList();
     }
 
@@ -53,6 +52,7 @@ public class NodeList {
     }
 
     public void remove(Object key) {
+
         if (this.heap.containsKey(key)) {
             this.heap.remove(key);
         } else {

### 2015-03-20 src/main/java/edu/colostate/cs/gc/process/MessageWorker.java

@@ -89,6 +89,5 @@ public class MessageWorker implements Runnable {
         }
 
         this.latch.countDown();
-
     }
 }

### 2015-03-20 src/main/java/edu/colostate/cs/gc/process/TopEventProcessor.java

@@ -27,10 +27,8 @@ public class TopEventProcessor extends TripProcessor {
 
     private List<NodeValue> lastRoutesList;
 
-    public static final int MAX_QUEUE_SIZE = 100;
-
     public TopEventProcessor(EventWriter eventWriter) {
-        this.nodeList = new NodeList();
+        this.nodeList = new NodeList(64);
         this.lastRoutesList = new ArrayList<NodeValue>();
         this.eventWriter = eventWriter;
 
@@ -58,23 +56,8 @@ public class TopEventProcessor extends TripProcessor {
     }
 
     public synchronized void processEvent(TripEvent event) {
+
         TopEvent topEvent = (TopEvent) event;
-        Queue<TopEvent> queue = this.queues.get(topEvent.getProcessorID());
-        if (queue.size() > MAX_QUEUE_SIZE){
-            // if it exceeds max size then we need to move the existing messages to application
-            while (queue.size() > MAX_QUEUE_SIZE){
-                // we know this queue should have at messages
-                int minIndex = topEvent.getProcessorID();
-                int minSeq = queue.peek().getSeqNo();
-                for (int i = 0; i < this.numbOfProcessors; i++) {
-                    if (!this.queues.get(i).isEmpty() && this.queues.get(i).peek().getSeqNo() < minSeq) {
-                        minIndex = i;
-                        minSeq = this.queues.get(i).peek().getSeqNo();
-                    }
-                }
-                processOrderedMessage(this.queues.get(minIndex).poll());
-            }
-        }
         this.queues.get(topEvent.getProcessorID()).add(topEvent);
 
         int minIndex = 0;

### 2015-03-20 src/main/java/edu/colostate/cs/gc/profit/ProfitCalculator.java

@@ -22,7 +22,7 @@ public class ProfitCalculator extends TripProcessor {
 
     private Queue<PaymentEvent> paymentWindow = new LinkedList<PaymentEvent>();
     private Queue<PaymentEvent> dropWindow = new LinkedList<PaymentEvent>();
-    private NodeList nodeList = new NodeList();
+    private NodeList nodeList = new NodeList(2048);
 
     private long startTime;
     private boolean isStarted;
@@ -192,7 +192,6 @@ public class ProfitCalculator extends TripProcessor {
                                               profitCellNode.getCell()));
             }
 
-
             this.processor.processEvent(topProfitsEvent);
             this.lastCellSet = currentCellSet;
         }
@@ -209,7 +208,6 @@ public class ProfitCalculator extends TripProcessor {
     }
 
     public void close() {
-
     }
 
 }

### 2015-03-20 src/main/java/edu/colostate/cs/gc/profit/ProfitEventEmitter.java

@@ -63,6 +63,7 @@ public class ProfitEventEmitter implements Adaptor {
             String line;
 
             int seqNo = 1;
+            int numberOfEvents = 0;
 
             try {
                 barrier.await();
@@ -103,6 +104,7 @@ public class ProfitEventEmitter implements Adaptor {
                         if (pickUpEvent.getFare() > 0){
                             bufferNumber = pickUpEvent.getPickUpCell().hashCode() % messageBuffers.length;
                             messageBuffers[bufferNumber].addMessage(pickUpEvent);
+                            numberOfEvents++;
 
                             PaymentEvent dropOffEvent = new PaymentEvent();
                             dropOffEvent.setSeqNo(seqNo);
@@ -143,6 +145,7 @@ public class ProfitEventEmitter implements Adaptor {
             System.out.println("Display statistics ...");
             System.out.println("Total time (ms) " + totalTime);
             System.out.println("Total time (Min) " + totalTime * 1.0 / 60000);
+            System.out.println("Throughput (Msg/s) " + numberOfEvents * 1000.0 / totalTime);
 
             bufferedReader.close();
 
@@ -175,7 +178,7 @@ public class ProfitEventEmitter implements Adaptor {
     }
 
     public static void main(String[] args) {
-        int numberOfBuffers = 4;
+        int numberOfBuffers = Integer.parseInt(args[1]);
 
         CyclicBarrier barrier = new CyclicBarrier(numberOfBuffers + 1);
         CountDownLatch latch = new CountDownLatch(numberOfBuffers);

### 2015-03-20 src/main/java/edu/colostate/cs/gc/route/RouteEventEmitter.java

@@ -43,6 +43,7 @@ public class RouteEventEmitter implements Adaptor {
 
             String line = null;
             int seqNo = 1;
+            int numberOfEvents = 0;
 
             try {
                 barrier.await();
@@ -75,6 +76,7 @@ public class RouteEventEmitter implements Adaptor {
                         dropOffEvent.setRoute(route);
                         int bufferNumber = route.hashCode() % messageBuffers.length;
                         messageBuffers[bufferNumber].addMessage(dropOffEvent);
+                        numberOfEvents++;
                     }
 
                 } catch (Exception e) {
@@ -97,6 +99,7 @@ public class RouteEventEmitter implements Adaptor {
             System.out.println("Display statistics ...");
             System.out.println("Total time (ms) " + totalTime);
             System.out.println("Total time (Min) " + totalTime * 1.0 / 60000 );
+            System.out.println("Throughput (Msg/s) " + numberOfEvents * 1000.0 / totalTime);
 
             bufferedReader.close();
 
@@ -149,7 +152,7 @@ public class RouteEventEmitter implements Adaptor {
 
 
     public static void main(String[] args) {
-        int numberOfBuffers = 3;
+        int numberOfBuffers = Integer.parseInt(args[1]);
 
         CyclicBarrier barrier = new CyclicBarrier(numberOfBuffers + 1);
         CountDownLatch latch = new CountDownLatch(numberOfBuffers);

### 2015-03-20 src/main/java/edu/colostate/cs/gc/route/RouteProcessor.java

@@ -24,7 +24,7 @@ public class RouteProcessor extends TripProcessor {
 
     private Queue<DropOffEvent> window = new LinkedList<DropOffEvent>();
 
-    NodeList nodeList = new NodeList();
+    NodeList nodeList = new NodeList(8192);
 
     private boolean isStarted = false;
     private long startTime;
@@ -140,6 +140,5 @@ public class RouteProcessor extends TripProcessor {
     }
 
     public void close() {
-
     }
 }

### 2015-03-20 src/test/java/edu/colostate/cs/gc/list/TestNodeList.java

@@ -15,7 +15,7 @@ public class TestNodeList extends TestCase {
 
     public void testRemoveNodes(){
 
-        NodeList nodeList = new NodeList();
+        NodeList nodeList = new NodeList(100);
 
         nodeList.add("route1", new CountNode(5, "route1"));
         assertTrue(nodeList.checkConsistency());
@@ -79,7 +79,7 @@ public class TestNodeList extends TestCase {
 
     public void testDecrementPosition(){
 
-        NodeList nodeList = new NodeList();
+        NodeList nodeList = new NodeList(100);
 
         nodeList.add("route1", new CountNode(5, "route1"));
         assertTrue(nodeList.checkConsistency());
@@ -145,7 +145,7 @@ public class TestNodeList extends TestCase {
     }
 
     public void testIncrementPosition(){
-        NodeList nodeList = new NodeList();
+        NodeList nodeList = new NodeList(100);
 
         nodeList.add("route1", new CountNode(5, "route1"));
         assertTrue(nodeList.checkConsistency());
@@ -229,7 +229,7 @@ public class TestNodeList extends TestCase {
     }
 
     public void testIncrementWithoutHeap(){
-        NodeList nodeList = new NodeList();
+        NodeList nodeList = new NodeList(100);
 
         nodeList.add("route1", new CountNode(5, "route1"));
         assertTrue(nodeList.checkConsistency());

### 2015-03-24 src/main/java/edu/colostate/cs/gc/process/MessageWorker.java
configuring time windows

@@ -89,5 +89,7 @@ public class MessageWorker implements Runnable {
         }
 
         this.latch.countDown();
+
+        this.processor.close();
     }
 }

### 2015-03-24 src/main/java/edu/colostate/cs/gc/profit/LocalProfitProcessor.java

@@ -33,9 +33,10 @@ public class LocalProfitProcessor extends TripProcessor {
     public void initialise(Container container, Map<String, String> parameters) {
         super.initialise(container, parameters);
         this.numOfProcessors = Integer.parseInt(parameters.get("processors"));
+        int windowSize = Integer.parseInt(parameters.get("windowSize"));
         this.processors = new ProfitCalculator[this.numOfProcessors];
         for (int i = 0; i < this.numOfProcessors; i++) {
-            this.processors[i] = new ProfitCalculator(this.processor, this.numOfProcessors);
+            this.processors[i] = new ProfitCalculator(this.processor, this.numOfProcessors, windowSize);
         }
     }
 }

### 2015-03-24 src/main/java/edu/colostate/cs/gc/profit/ProfitCalculator.java

@@ -31,16 +31,18 @@ public class ProfitCalculator extends TripProcessor {
     private List<NodeValue> lastCellList;
 
     private int numOfProcessors;
+    private int windowSize;
 
     public ProfitCalculator() {
         this.lastCellList = new ArrayList<NodeValue>();
         this.lastCellSet = new HashSet<Cell>();
     }
 
-    public ProfitCalculator(TripProcessor tripProcessor, int numOfProcessors) {
+    public ProfitCalculator(TripProcessor tripProcessor, int numOfProcessors, int windowSize) {
         this();
         this.processor = tripProcessor;
         this.numOfProcessors = numOfProcessors;
+        this.windowSize = windowSize * Constants.SMALL_WINDOW_SIZE;
     }
 
     @Override
@@ -93,7 +95,7 @@ public class ProfitCalculator extends TripProcessor {
             }
 
             while ((this.paymentWindow.size() > 0) &&
-                    (this.paymentWindow.peek().isExpired(paymentEvent.getDropOffTimeMillis(), Constants.SMALL_WINDOW_SIZE))) {
+                    (this.paymentWindow.peek().isExpired(paymentEvent.getDropOffTimeMillis(), this.windowSize))) {
                 PaymentEvent expiredEvent = this.paymentWindow.poll();
                 ProfitCellNode profitCellNode = (ProfitCellNode) this.nodeList.get(expiredEvent.getPickUpCell());
                 double preProfitability = profitCellNode.getProfitability();
@@ -133,7 +135,7 @@ public class ProfitCalculator extends TripProcessor {
             }
 
             while ((this.dropWindow.size() > 0)
-                    && (this.dropWindow.peek().isExpired(paymentEvent.getDropOffTimeMillis(), Constants.LARGE_WINDOW_SIZE))) {
+                    && (this.dropWindow.peek().isExpired(paymentEvent.getDropOffTimeMillis(), this.windowSize * 2))) {
                 PaymentEvent expiredEvent = this.dropWindow.poll();
 
                 if (this.nodeList.containsKey(expiredEvent.getDropOffCell())) {
@@ -196,7 +198,6 @@ public class ProfitCalculator extends TripProcessor {
             this.lastCellSet = currentCellSet;
         }
         this.lastCellList = currentList;
-
     }
 
     private Set<Cell> getCellSet(List<NodeValue> cellList) {

### 2015-03-24 src/main/java/edu/colostate/cs/gc/profit/ProfitCellNode.java

@@ -2,6 +2,9 @@ package edu.colostate.cs.gc.profit;
 
 import edu.colostate.cs.gc.event.Cell;
 import edu.colostate.cs.gc.list.NodeValue;
+import edu.colostate.cs.gc.profit.median.DynamicFieldMF;
+import edu.colostate.cs.gc.profit.median.MedianFinder;
+import edu.colostate.cs.gc.profit.median.PriorityQueueMF;
 
 import java.util.Comparator;
 import java.util.HashSet;
@@ -30,6 +33,9 @@ public class ProfitCellNode implements NodeValue {
 
     private int seqNo;
 
+    private MedianFinder medianFinder = new PriorityQueueMF();
+//    private MedianFinder medianFinder = new DynamicFieldMF();
+
     /**
      * this constructor is being created from an drop off event.
      */
@@ -52,6 +58,7 @@ public class ProfitCellNode implements NodeValue {
     public ProfitCellNode(double fare, Cell cell, int seqNo) {
         this.midFare = fare;
         this.numOfFares = 1;
+        this.medianFinder.add(fare);
         this.numOfEmptyTaxis = 0;
         this.profitability = UNDEFINED_PROFITABILITY;
         this.cell = cell;
@@ -67,19 +74,6 @@ public class ProfitCellNode implements NodeValue {
         this.profitability = newValue.profitability;
         this.seqNo = newValue.seqNo;
     }
-
-    private PriorityQueue<Double> lowQueue = new PriorityQueue<Double>(20, new Comparator<Double>() {
-
-        public int compare(Double o1, Double o2) {
-            return o2.compareTo(o1);
-        }
-    });
-    private PriorityQueue<Double> highQueue = new PriorityQueue<Double>(20, new Comparator<Double>() {
-
-        public int compare(Double o1, Double o2) {
-            return o1.compareTo(o2);
-        }
-    });
 
     public void reduceEmptyTaxi() {
         if (this.numOfEmptyTaxis > 0) {
@@ -99,28 +93,8 @@ public class ProfitCellNode implements NodeValue {
         // this means this taxi has leaved from this location. i.e we need to reduce the empty taxis.
         this.numOfFares++;
 
-        if (this.numOfFares % 2 == 0) {
-            if (fare <= this.midFare) {
-                this.lowQueue.add(fare);
-                double highOfLow = this.lowQueue.peek();
-                this.highQueue.add(this.midFare);
-                this.midFare = (this.midFare + highOfLow) / 2;
[…]

### 2015-03-24 src/main/java/edu/colostate/cs/gc/profit/ProfitEventEmitter.java

@@ -179,6 +179,7 @@ public class ProfitEventEmitter implements Adaptor {
 
     public static void main(String[] args) {
         int numberOfBuffers = Integer.parseInt(args[1]);
+        int windowSize = Integer.parseInt(args[2]);
 
         CyclicBarrier barrier = new CyclicBarrier(numberOfBuffers + 1);
         CountDownLatch latch = new CountDownLatch(numberOfBuffers);
@@ -187,7 +188,8 @@ public class ProfitEventEmitter implements Adaptor {
         //initialize buffers
         MessageBuffer[] messageBuffers = new MessageBuffer[numberOfBuffers];
         for (int i = 0; i < messageBuffers.length; i++) {
-            messageBuffers[i] = new MessageBuffer(new ProfitCalculator(topProfitProcessor, numberOfBuffers), barrier, latch);
+            messageBuffers[i] = new MessageBuffer(
+                                new ProfitCalculator(topProfitProcessor, numberOfBuffers, windowSize), barrier, latch);
         }
         new ProfitEventEmitter().loadData(args[0], messageBuffers, barrier, latch);
         topProfitProcessor.close();

### 2015-03-24 src/main/java/edu/colostate/cs/gc/profit/median/DynamicFieldMF.java

@@ -1,0 +1,198 @@
+package edu.colostate.cs.gc.profit.median;
+
+import edu.colostate.cs.gc.list.DynamicHeap;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/18/15
+ * Time: 9:50 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class DynamicFieldMF extends MedianFinder {
+
+    private DynamicHeap minHeap = new DynamicHeap(8);
+    private DynamicHeap maxHeap = new DynamicHeap(8);
+
+    // this value is equals to twice of the median field size
+    private int medianCount;
+
+    // position of the median field
+    private int medianPosition;
+
+    // this means median is in middle of two buckets. so real median will be the middle value.
+    private boolean isInMiddle;
+
+    public DynamicFieldMF() {
+        this.median = 0;
+        this.isInMiddle = true;
+    }
+
+    public void add(double value) {
+        if (this.isInMiddle) {
+            // in this case we don't care about the equality since the median is a fake value. and there is no
+            // possibility that both are equal since we have unique buckets.
+            if (value > this.median) {
+                addToMaxHeap(value);
+                // new median will be the minimum value of this map
+                pullMedianFromMaxHeap();
+            } else {
+                addToMinHeap(value);
+                pullMedianFromMinHeap();
+            }
+            this.isInMiddle = false;
+        } else {
+            // i.e there is a median class available.
+            // when adding an element we assume it adds to end of the bucket. so need to remove from end as well.
+            if (value == this.median) {
+                this.medianCount = this.medianCount + 2;
+                this.medianPosition++;
+            } else if (value > this.median) {
+                addToMaxHeap(value);
+                // median position should advance by one.
+                this.medianPosition++;
+                if (this.medianPosition > this.medianCount) {
+                    //i.e median should in between.
+                    pushMedianToMinHeap();
+                    this.isInMiddle = true;
+                }
+            } else {
+                // i.e current value is less than median.
+                addToMinHeap(value);
+                this.medianPosition--;
+                if (this.medianPosition < 2) {
+                    // i.e median should be in between
+                    pushMedianToMaxHeap();
+                    this.isInMiddle = true;
+                }
+            }
+        }
+
+    }
+
[…]

### 2015-03-24 src/main/java/edu/colostate/cs/gc/profit/median/MedianFinder.java

@@ -1,203 +1,22 @@
 package edu.colostate.cs.gc.profit.median;
 
-import edu.colostate.cs.gc.list.DynamicHeap;
-
 /**
  * Created with IntelliJ IDEA.
  * User: amila
- * Date: 3/18/15
- * Time: 9:50 AM
+ * Date: 3/23/15
+ * Time: 10:29 AM
  * To change this template use File | Settings | File Templates.
  */
-public class MedianFinder {
+public abstract class MedianFinder {
 
-    private DynamicHeap minHeap = new DynamicHeap(8);
-    private DynamicHeap maxHeap = new DynamicHeap(8);
+    protected double median;
 
-    private double median;
-    // this value is equals to twice of the median field size
-    private int medianCount;
+    public abstract void add(double value);
 
-    // position of the median field
-    private int medianPosition;
-
-    // this means median is in middle of two buckets. so real median will be the middle value.
-    private boolean isInMiddle;
-
-    public MedianFinder() {
-        this.median = 0;
-        this.isInMiddle = true;
-    }
+    public abstract void remove(double value);
 
     public double getMedian() {
-        return this.median;
+        return median;
     }
-
-    public void add(double value) {
-        if (this.isInMiddle) {
-            // in this case we don't care about the equality since the median is a fake value. and there is no
-            // possibility that both are equal since we have unique buckets.
-            if (value > this.median) {
-                addToMaxHeap(value);
-                // new median will be the minimum value of this map
-                pullMedianFromMaxHeap();
-            } else {
-                addToMinHeap(value);
-                pullMedianFromMinHeap();
-            }
-            this.isInMiddle = false;
-        } else {
-            // i.e there is a median class available.
-            // when adding an element we assume it adds to end of the bucket. so need to remove from end as well.
-            if (value == this.median) {
-                this.medianCount = this.medianCount + 2;
-                this.medianPosition++;
-            } else if (value > this.median) {
-                addToMaxHeap(value);
-                // median position should advance by one.
-                this.medianPosition++;
-                if (this.medianPosition > this.medianCount) {
-                    //i.e median should in between.
-                    pushMedianToMinHeap();
-                    this.isInMiddle = true;
-                }
-            } else {
-                // i.e current value is less than median.
-                addToMinHeap(value);
[…]

### 2015-03-24 src/main/java/edu/colostate/cs/gc/profit/median/PriorityQueueMF.java

@@ -1,0 +1,93 @@
+package edu.colostate.cs.gc.profit.median;
+
+import java.util.Comparator;
+import java.util.PriorityQueue;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/23/15
+ * Time: 10:33 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class PriorityQueueMF extends MedianFinder {
+
+    private PriorityQueue<Double> lowQueue = new PriorityQueue<Double>(20, new Comparator<Double>() {
+
+        public int compare(Double o1, Double o2) {
+            return o2.compareTo(o1);
+        }
+    });
+    private PriorityQueue<Double> highQueue = new PriorityQueue<Double>(20, new Comparator<Double>() {
+
+        public int compare(Double o1, Double o2) {
+            return o1.compareTo(o2);
+        }
+    });
+
+    private int numOfFares;
+
+    public PriorityQueueMF() {
+        this.numOfFares = 0;
+        this.median = 0;
+    }
+
+    @Override
+    public void add(double fare) {
+        this.numOfFares++;
+
+        if (this.numOfFares % 2 == 0) {
+            if (fare <= this.median) {
+                this.lowQueue.add(fare);
+                double highOfLow = this.lowQueue.peek();
+                this.highQueue.add(this.median);
+                this.median = (this.median + highOfLow) / 2;
+            } else {
+                this.highQueue.add(fare);
+                double lowOfHigh = this.highQueue.peek();
+                this.lowQueue.add(this.median);
+                this.median = (this.median + lowOfHigh) / 2;
+            }
+        } else {
+            // now existing median is a fake median which is equals to mid point of two medians
+            if (fare <= this.median) {
+                this.lowQueue.add(fare);
+                this.median = this.lowQueue.poll();
+            } else {
+                this.highQueue.add(fare);
+                this.median = this.highQueue.poll();
+            }
+        }
+    }
+
+    @Override
+    public void remove(double fare) {
+        this.numOfFares--;
+
+        if (this.numOfFares == 0) {
+            this.median = 0;
+        } else {
+            if (this.numOfFares % 2 == 0) {
+                if (fare == this.median) {
+                    this.median = (this.lowQueue.peek() + this.highQueue.peek()) / 2;
+                } else if (fare > this.median) {
+                    this.highQueue.remove(fare);
+                    this.highQueue.add(this.median);
+                    this.median = (this.median + this.lowQueue.peek()) / 2;
+                } else {
[…]

### 2015-03-24 src/main/java/edu/colostate/cs/gc/util/Constants.java

@@ -9,8 +9,8 @@ package edu.colostate.cs.gc.util;
  */
 public interface Constants {
 
-    public static final int LARGE_WINDOW_SIZE = 30 * 60 * 1000;
     public static final int SMALL_WINDOW_SIZE = 15 * 60 * 1000;
+    public static final int LARGE_WINDOW_SIZE = SMALL_WINDOW_SIZE * 2;
 
     public static final double EAST_CELL_SIZE_500 = 0.005986;
     public static final double SOUTH_CELL_SIZE_500 = 0.004491556;

### 2015-03-24 src/test/java/edu/colostate/cs/gc/profit/median/TestDynamicHeapMF.java

@@ -9,11 +9,11 @@ import junit.framework.TestCase;
  * Time: 11:09 AM
  * To change this template use File | Settings | File Templates.
  */
-public class TestMedianFinder extends TestCase {
+public class TestDynamicHeapMF extends TestCase {
 
     public void testAddUnique(){
 
-        MedianFinder medianFinder = new MedianFinder();
+        DynamicFieldMF medianFinder = new DynamicFieldMF();
         medianFinder.add(20.0);
         assertEquals(medianFinder.getMedian(), 20.0);
 
@@ -32,7 +32,7 @@ public class TestMedianFinder extends TestCase {
     }
 
     public void testAddDuplicate(){
-        MedianFinder medianFinder = new MedianFinder();
+        DynamicFieldMF medianFinder = new DynamicFieldMF();
         medianFinder.add(25.0);
         assertEquals(medianFinder.getMedian(), 25.0);
         medianFinder.add(25.0);
@@ -83,7 +83,7 @@ public class TestMedianFinder extends TestCase {
     }
 
     public void testRemoveUnique(){
-        MedianFinder medianFinder = new MedianFinder();
+        DynamicFieldMF medianFinder = new DynamicFieldMF();
         medianFinder.add(20.0);
         medianFinder.add(25.0);
         medianFinder.add(17.0);
@@ -111,7 +111,7 @@ public class TestMedianFinder extends TestCase {
     }
 
     public void testRemoveDuplicate(){
-        MedianFinder medianFinder = new MedianFinder();
+        DynamicFieldMF medianFinder = new DynamicFieldMF();
         medianFinder.add(25.0);
         medianFinder.add(25.0);
 

### 2015-03-24 src/main/java/edu/colostate/cs/gc/GCEventEmitter.java
runing queries together

@@ -1,0 +1,227 @@
+package edu.colostate.cs.gc;
+
+import edu.colostate.cs.gc.event.Cell;
+import edu.colostate.cs.gc.event.DropOffEvent;
+import edu.colostate.cs.gc.event.PaymentEvent;
+import edu.colostate.cs.gc.event.Route;
+import edu.colostate.cs.gc.process.MessageBuffer;
+import edu.colostate.cs.gc.profit.ProfitCalculator;
+import edu.colostate.cs.gc.profit.TopProfitProcessor;
+import edu.colostate.cs.gc.route.RouteProcessor;
+import edu.colostate.cs.gc.route.TopRouteProcessor;
+import edu.colostate.cs.gc.util.Constants;
+import edu.colostate.cs.gc.util.Util;
+
+import java.io.BufferedReader;
+import java.io.FileNotFoundException;
+import java.io.FileReader;
+import java.io.IOException;
+import java.util.concurrent.BrokenBarrierException;
+import java.util.concurrent.CountDownLatch;
+import java.util.concurrent.CyclicBarrier;
+
+/**
+ * Created with IntelliJ IDEA.
+ * User: amila
+ * Date: 3/24/15
+ * Time: 9:21 AM
+ * To change this template use File | Settings | File Templates.
+ */
+public class GCEventEmitter {
+
+    private void loadData(String fileName,
+                          MessageBuffer[] routeBuffers,
+                          MessageBuffer[] profitBuffers,
+                          CyclicBarrier barrier,
+                          CountDownLatch latch) {
+
+        try {
+
+            BufferedReader bufferedReader = new BufferedReader(new FileReader(fileName));
+            String line;
+
+            int seqNo = 1;
+
+            try {
+                barrier.await();
+            } catch (InterruptedException e) {
+                e.printStackTrace();  //To change body of catch statement use File | Settings | File Templates.
+            } catch (BrokenBarrierException e) {
+                e.printStackTrace();  //To change body of catch statement use File | Settings | File Templates.
+            }
+
+            System.out.println("Start processing messages ...");
+            long currentTime = System.currentTimeMillis();
+
+
+            while ((line = bufferedReader.readLine()) != null) {
+
+                try {
+                    String[] values = line.split(",");
+
+                    // first process the longitude and latitude coordinates.
+                    double pickUpLongitude = Double.parseDouble(values[6]);
+                    double pickUpLatitude = Double.parseDouble(values[7]);
[…]

### 2015-03-24 paper/experiments.tex
fixing typos

@@
 We conduced several experiments to measure the soundness of our solution.
-Since the parallel message processing gives the better performance, we present details of those experiments with distributed setup as well a singel JVM setup.
+Since the parallel message processing gives the better performance, we present details of those experiments with distributed setup as well a single JVM setup.
 All these experiments were performed on a LAN with a network bandwidth of 1Gbps and all machines were Intel(R) Xeon(R) 2.4GHz, 4 core duo machines with 16 GB of memory.
@@
 We deployed this process graph to a distributed stream processing system we have developed, and  measured the execution time and the average message latency varying the number of route processors with the data set given (173 million records).
-In local set up, we use a new thread to execute a new instance and in distributed setup we add a new node to execute the new instance.
+In the local set up, we use a new thread to execute a new instance and in the distributed setup we add a new node to execute the new instance.
 
@@
 Figure \ref{latency_route} shows the latencies.
-For single JVM setup latency get reduced at 2 and slightly get increase due to context switching as in the execution times.
+For single JVM setup latency get reduced at 2 and slightly increases due to context switching as in the execution times.
 For distributed set up initial latency is high due to communication and message serialization parsing overheads.
-With higher number of processors latency get reduced.
+With higher number of processors latency gets reduced.
 
@@
 \textit{ProfitEventEmitter} reads the data file and calculates the pickup cell and dropoff cell details using longitude and latitude coordinates.
-Then it creates two events one for dropoff part and other to pickup part, with different sequence numbers and distributes the events among \textit{ProfictCalculators} using the hash value of the cell.
+Then it creates two events one for dropoff part and other to pickup part, with different sequence numbers and distributes the events among \textit{ProfitCalculators} using the hash value of the cell.
 In this way cells are partitioned into different \textit{ProfitCalculator} instances and each instance gets all the events relevant to its' cell set.
 We use a set of threads (equals to the number of Profict Calculators) to emit data as in the previous query.
-Finally \textit{TopProfictProcessor} aggregates events and writes to a file.
+Finally \textit{TopProfitProcessor} aggregates events and writes to a file.
 Similar to earlier experiment,  we deployed this process graph to our system and measured the execution time and the average message latency with the data set given for both single JVM setup and distributed setup.

### 2015-03-24 paper/paper.tex
updated the paper
fixing typos

@@
 Processing complex queries on unbounded event streams in real time, is a challenge for many organizational data processing systems.
-These systems are expected to process data with minimal latency to generate real time events, and high through put to minimize the required hardware.
-In this regard DEBS Grand Challenge 2015 focus on evaluating two queries (frequent routes and  profitable cells) in real time with minimal latency and high throughput.
+These systems are expected to process data with reduced latency to generate real time events, and high throughput to minimize the required hardware.
+In this regard, DEBS Grand Challenge 2015 focuses on evaluating two queries (frequent routes and  profitable cells) in real time with low latency and high throughput.
 These queries involve processing windows of thousands of records.
-Firstly, such processing demands efficient data structures and algorithms to minimize the processing over head.
-Secondly, system should partition data to process them in parallel to make system scalable.
+Firstly, such processing demands efficient data structures and algorithms to minimize the processing overhead.
+Secondly, the system should partition data to process them in parallel to make it scalable.
 
-In this paper, we present a set of data structures which can be used to evaluate two queries with O(log(n)) time complexity and a data partition technique to process events in parallel.
-Then we evaluate our solution within signal node as well as in distributed setups.
-We could process frequent routes query with 173 million dataset within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
+In this paper, we present a set of data structures which can be used to evaluate the aforementioned queries with O(log(n)) time complexity and a data partitioning technique to process events in parallel.
+Then we evaluate our solution within signal node as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN.
+We could process frequent routes query with 173 million data set within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
 \end{abstract}
@@
 \section{Introduction}
-introduction
+%introduction
 

### 2015-03-24 paper/query_evaluation.tex
fixing typos

@@
 This sequence number is used to order \textit{QueryEvaluator} results.
-For real implementation, each query has its own \textit{EventEmitter} which reads each line of the data file and generates an event with relevant information.
+For real implementation, each query has its own \textit{EventEmitter} that reads each line of the data file and generates an event with the relevant information.
 \textit{QueryEvaluator} evaluates the query against the received event and generates top ten value changed event.
-In general we can evaluate queries sequentially and parallel.
-In a sequential execution, all events goes to one \textit{QueryEvaluator} instance and in parallel execution, event stream is partitioned using the key field and evaluate simultaneously with different \textit{QueryEvaluator} instances.
+In general we can evaluate queries sequentially and in parallel.
+In a sequential execution, all events goes to one \textit{QueryEvaluator} instance and in a parallel execution, the event stream is partitioned using the key field and evaluate simultaneously with different \textit{QueryEvaluator} instances.
 
@@
 We can sequentially execute events either in a single Java Virtual Machine (JVM) or different JVMs (either in one machine or different machines) using TCP sockets to communicate among process instances.
-Figure \ref{sequential} shows such possible executions.
+Figure \ref{sequential} shows possible executions.
 
@@
 First both processors can be executed in same JVM with same thread.
-In that case, both processes executes sequentially.
+In that case, both processes execute sequentially.
 We can make two processes parallel by executing them with two threads and communicate using a queue in same JVM or communicate using a TCP connection for different JVMs.
@@
 \subsection{Parallel evaluation}
-The general way to parallelize event processing, is to partition the events according to key and run multiple instances to process data.
-Therefore in this section, we explain how to parallelize the query evaluation process taking frequent rote query into consideration.
+The general way to parallelize event processing, is to partition the events according to a key and run multiple instances to process data.
+Therefore in this section, we explain how to parallelize the query evaluation process taking frequent route query into consideration.
 Same concepts can be applied to profitable cell evaluation as well.
@@
 For frequent route query, route events can be partitioned using the hash code of the route object and process different route sets in different \textit{QueryEvaluators}.
-In such a setup, each \textit{QueryEvaluator} generates an event when top ten routes get changed for its route set.
+In such a setup, each \textit{QueryEvaluator} generates an event when the top ten routes get changed for its route set.
 Since this is not the desired result, we can add another process to aggregate these top result events and generate global route change events.
@@
 \textit{ResultAggregator} keeps a set of route counts received by \textit{QueryEvaluators}.
-Upon receiving a new event, it can update the existing route count values and check whether global top ten route counts has been changed.
+Upon receiving a new event, it can update the existing route count values and check whether the global top ten route counts have changed.
 If there is a change, it can generate a top ten route changed event.
@@
 The top ten frequent routes can be changed due to these expiration events as well.
-Lets say in one \textit{QuerayEvaluator} case, a particular event cause two routes R1 and R2 to change and generates a top ten route change event.
+Lets say in one \textit{QuerayEvaluator} case, a particular event causes two routes R1 and R2 to change and generates a top ten route change event.
 In the distributed setup these two routes can be in two \textit{QueryEvaluators} and that can cause two top ten route change events.
@@
 First it removes the expired keys.
-Then it update the existing values or add new values and move the position of the list accordingly.
+Then it updates the existing values or adds new values and moves the position in the list accordingly.
 Finally it generates an event if there is a change to the top list.

### 2015-03-24 paper/query_implementation.tex
fixing typos

@@
 The 2015 Grand Challenge problem has two queries.
-First is to evaluate top ten most frequent routes within last 30 minutes and second is to evaluate top ten profitable areas within last 15 minutes.
-We can use a composite object to keep route details for each route and fare details to each cell to compare values among routes and cells.
+First is to evaluate top ten most frequent routes within the last 30 minutes and second is to evaluate top ten profitable areas within the last 15 minutes.
+We use a composite object to manage details for each route and fare details to each cell to compare values among routes and cells.
 However, each time window can have thousands of route and cell details.
-Therefore an efficient data structure is required to add, remove, update and retrieve top ten values from a set of objects for a given time window efficiently.
-Further since each event is received with route and cell details, above mentioned operations should perform using route and cell details as the key.
-In order to support all these functions with O(log(n)) time complexity we developed a data structure called \textit{NodeList}.
-Following parts of this section describes how we have achieve this time complexity with \textit{NodeList} data structure and how we have implemented the queries using that data structure.
+Therefore, an efficient data structure is required to add, remove, update and retrieve top ten values from a set of objects for a given time window efficiently.
+Further, since each event is received with route and cell details, the aforementioned operations should be performed using route and cell details as the key.
+In order to support all these functions with O(log(n)) time complexity we have developed a data structure called \textit{NodeList}.
+The remaining of this section describes how we have achieve this time complexity with the \textit{NodeList} data structure and how we have implemented the queries using that data structure.
 
@@
 The \textit{NodeValue} is an interface to plug any user defined type as a value.
-For an example, the route query can use a \textit{RouteCount} object to keep route count while profitable cells query can use \textit{CellProfict} object to keep fare details.
+For an example, the route query can use a \textit{RouteCount} object to keep route count while profitable cells query can use \textit{CellProfit} object to keep track of fare details.
 This interface contains a method called compare to define the order of the values.
@@
 The \texttt{decrementPostion} operation moves the corresponding value towards position 0 until the correct position according to sorted order.
-Therefore this method should be called after increasing a value of an existing object.
+Therefore this method should be called after increasing the value of an existing object.
 Similarly \texttt{incrementPosition} operation can be used to move a value to correct position after decreasing the value of an existing object.
 \texttt{GetTopValues} operation returns the top ten values of the list.
-Query algorithm sections further elaborate usage of these methods.
+The query algorithm sections further elaborate usage of these methods.
 
 One way of implementing this data structure is to use a heap.
-Heap supports adding and extracting maximum value operations in O(log(n)) time.
+A heap supports adding and extracting maximum value operations in O(log(n)) time.
 However since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back.
@@
 
-\textit{NodeList} data structure uses two data structures called \textit{OrderedList} to keep first 10 values and \textit{DynamicHeap} to keep other values.
+The \textit{NodeList} data structure uses two data structures called \textit{OrderedList} to keep first 10 values and \textit{DynamicHeap} to keep other values.
 In this way it has to change only the last value of \textit{OrderedList} and maximum value of \textit{DynamicHeap} as required.
@@
 Table \ref{orderedlist_api} shows the operations of the \textit{OrderedList} data structure.
-First six methods has the same meanings as the \textit{NodeList} data structure and last two can be used to get the size of the list and the value of the last object.
+First six methods has the same meaning as the \textit{NodeList} data structure and last two can be used to get the size of the list and the value of the last object.
 We internally keep a counter and a pointer to tail to support these operations in O(1) time.
@@
 Figure \ref{dynamicheap_impl} shows the implementation details.
-Conventional heaps only support \texttt{add} and \texttt{extractMax} operations since heap operations requires to know the element position in heap.
-When adding an element, we add it to last position and move that up and extracting maximum element start at array position 0.
+Conventional heaps only support \texttt{add} and \texttt{extractMax} operations since heap operations require knowing the position of the element in underlying array.
 In order to support \texttt{moveUp}, \texttt{moveDown} and \texttt{remove} operations, we keep the array index with the object itself.
+As shown in figure \ref{dynamicheap_impl} \textit{HeapNode} object contains the array index and the \textit{NodeValue} given by the user.
 This allows us to retrieve array index in O(1) time and move the element O(log(n)) time.
@@
 
-Frequent route query requires to find the top frequent routes within last 30 minutes.
+The frequent route query entails finding the most frequent routes within the last 30 minutes.
 Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.
-Since queue operations are supported in O(1) time, query evaluation happens in O(log(n)) time.
+Since queue operations are supported in O(1) time, query evaluations happen in O(log(n)) time.
 Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to process an event.
-This algorithm assumes there is a global \textit{NodeList} object and route counts stored in an Object called \textit{RouteCount}.
+This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an Object called \textit{RouteCount}.
 
@@
 Processing an event can cause add, remove or update values of the existing time window objects.
-Therefore query processing mainly involves with performing above operations to the underlying \textit{NodeList} data structure and compare before and after top ten routes.
+Therefore query processing mainly involves performing the above operations with the underlying \textit{NodeList} data structure and comparing top ten routes before and after.
 As shown in Algorithm \ref{frequent_route_algorithm}, when an event arrives, first it puts the event to time window to keep track for last 30 minutes.
@@
 
-We have implemented profitable cells query in a similar manner to the frequent route query using  time windows. Complexity of the query raised following issues and we have solved them as given bellow.
+We have implemented profitable cells query in a similar manner to the frequent route query using  time windows. Complexity of the query raised following issues and we have solved them as given below.
 \begin{enumerate}
@@
 However the \texttt{delete} operation of \textit{java.util.PriorityQueue} takes O(n) time, since it traverse along the array to find the index of the object.
-We tried with our \textit{DynamicHeap} implementation to fix this.
+We implemented a similar algorithm using our \textit{DynamicHeap} to keep higher and lower values than median with value count.
+In worst case it takes O(log(n)) time to add or remove a value, in best case (i.e if the value is already present and that does not cause median to change) it takes O(1) time.
 But that did not improved the performance.
-Firstly, these heap sizes does not grow over the value of 145 and there are same values repeating in the list increasing the probability of find a value without traversing the whole list.
+First, these heap sizes do not grow over 145 and there are same values repeating in the list increasing the probability of finding a value without traversing the whole list.
 Secondly, our implementation create more objects compared to just keeping double values though the time complexity is O(log(n)).
-	\item In order to keep tract of empty taxis, we need to process both pickup cell and dropoff cell details for each message.
-One way of doing this is to send all events for same process instance and process both details with one event.
-However this avoid the possibility of partitioning data across different processes.
+	\item In order to keep track of empty taxis, we need to process both pickup cell and dropoff cell details for each message.
+One way of doing this is to send all events for the same process instance and process both details with one event.
+However this avoids the possibility of partitioning data across different processes.
 We generate two events, one for pickup data and one for dropoff data to make the process parallel.
@@
 If a pickup event occurs before dropoff event expiration, then we remove that taxi id from that cell and reduce the empty taxi count.
-In such a case dropoff event expiration does not cause a reduction of taxi count.
+In such a case, dropoff event expiration does not cause a reduction of taxi count.
 \end{enumerate}

### 2015-03-24 paper/related_work.tex
fixing typos

@@
 Many such systems have been developed with different feature sets according to the applications they use.
-Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track the trends in web queries.
-All these system are based on Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data.
+Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track trends in web queries.
+All these systems are based on the Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data.
 Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence dynamic membership management.
-Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets RDDs \cite{zaharia2012resilient} and Muppet \cite{lam2012muppet} provides state persistence using key value store based on Cassandra \cite{lakshman2010cassandra}.
+Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets RDDs \cite{zaharia2012resilient} and Muppet \cite{lam2012muppet} provides state persistence using a \textless key, value\textgreater store based on Cassandra \cite{lakshman2010cassandra}.
 

### 2015-03-25 paper/experiments.tex
adding a new experiment

@@
 On the other hand in a real time system, this event ordering many not required since event processing happens at real time.
+
+\subsection{Scalability with Window Size}
+
+The grand challenge problem uses 15 minute window size for profitable cells query.
+However in a real world problems window sizes may required.
+Therefore finally we evaluated scalability of our solution with the window size.
+For this experiment, we measured the throughput of the profitable cells query using a single JVM and three instances by varying the window size.
+Further we compared the throughput with our \textit{DynamicHeap} based solution which is not effective in lower window sizes as well.
+
+\begin{figure}[!t]
+        \centering
+        \includegraphics[width=3.0in]{window.png}
+        \caption{Throughput variation with the window size.}
+        \label{window}
+\end{figure}
+
+Figure \ref{window} shows the results.
+For both solutions, throughput does not drop by half, even though the window size increases 64 times.
+And for large windows \textit{DynamicHeap} performs better than the \textit{PriorityQueue} based one.
+This is due to two reasons.
+Firstly, all operations of \textit{DynamicHeap} are O(log(n)).
+Secondly, out of 173 million records there are only around 18000 unique fares.
+Therefore \textit{DyamicHeap} based algorithm does not consume memory beyond that.
+However , since \textit{PriortyQueue} based implementation keeps all values in memory it consumes more memory for large windows.

### 2015-03-25 paper/experiments.tex
re phasing experiments

@@
 \section{Experiments}
-We conduced several experiments to measure the soundness of our solution.
-Since the parallel message processing gives the better performance, we present details of those experiments with distributed setup as well a single JVM setup.
+We conducted several experiments to study the scalability of our solution.
+Firstly we study the throughput and mean message latency variation with the number of \textit{QueryEvaluators} for both single JVM and distributed setups.
+In an ideal system, we should observe linear throughput increment with a constant message latency.
+However there are some exceptions and we have provided our explanations on them.
+Finally we study the throughput variation with the window size to study the suitability of our solution for higher window sizes.
 All these experiments were performed on a LAN with a network bandwidth of 1Gbps and all machines were Intel(R) Xeon(R) 2.4GHz, 4 core duo machines with 16 GB of memory.
@@
 \textit{RouteEmitter} contains a set of threads (equals to the number of \textit{RouteProcessors}) to emit events more efficiently.
-When processing each line first it reads the data line and creates a Route object using longitude and latitude coordinates of pickup and dropoff locations.
-Then it distributes the events among the \textit{RouteProcessors} using the hash value of the Route object.
+When processing each line first it reads the data line and creates a \textit{Route} object using longitude and latitude coordinates of pickup and dropoff locations.
+Then it distributes the events among the \textit{RouteProcessors} using the hash value of the \textit{Route} object.
 \textit{TopRouteProcessor} aggregates the results and saves them to a file.
 
-We deployed this process graph to a distributed stream processing system we have developed, and  measured the execution time and the average message latency varying the number of route processors with the data set given (173 million records).
+We deployed this process graph to a distributed stream processing system we have developed, and  measured the throughput and the mean message latency varying the number of \textit{RouteProcessors} with the data set given (173 million records).
 In the local set up, we use a new thread to execute a new instance and in the distributed setup we add a new node to execute the new instance.
@@
         \centering
-        \includegraphics[width=3.0in]{time_route.png}
-        \caption{Execution times to evaluate top ten frequent routes.}
-        \label{time_route}
+        \includegraphics[width=3.0in]{throughput_route.png}
+        \caption{Throughput variation with the number of \textit{RouteProcessor} instances.}
+        \label{throughput_route}
 \end{figure}
 
-Figure \ref{time_route} shows the execution times.
-For both setups execution time is higher for single instance due to lack of parallelism of the system.
-For distributed setup time get reduced with addition of new nodes and comes to a constant level.
-This means our solution scales up to 4 node level and maintains a constant throughput.
-This is due to two reasons.
+Figure \ref{throughput_route} shows the throughput variation.
+For single JVM setup throughput increases from one instance to two instances.
+This is due to higher parallelism system achieves with two threads.
+After that throughput slightly decreases due to higher context switching with many threads.
+Similarly for distributed setup initially throughput increases with the number of   \textit{RouteProcessors} and achieves a steady limit.
+We observe two reasons for this non scalability after initial stages.
 Firstly, \textit{RouteProcessor} is not a computing intensive process.
 Secondly, there is only one \textit{RouteEventEmitter} and the speed at which it can read and send messages becomes the bottleneck.
-For single JVM set up highest throughput is achieved with two route processors.
-After that point time get slightly increases due to context switching.
+Initial throughput for distributed setup is less, due to message serialization, message deserialization and tcp overheads.
 
@@
         \includegraphics[width=3.0in]{latency_route.png}
-        \caption{Latencies to evaluate top ten frequent routes.}
+        \caption{Mean latency variation with the number of \textit{RouteProcessor} instances.}
         \label{latency_route}
@@
 
-Figure \ref{latency_route} shows the latencies.
-For single JVM setup latency get reduced at 2 and slightly increases due to context switching as in the execution times.
-For distributed set up initial latency is high due to communication and message serialization parsing overheads.
-With higher number of processors latency gets reduced.
+Figure \ref{latency_route} shows the latency variation.
+For single JVM setup latency get reduced at 2 and slightly increases after that due to context switching overhead with higher number of threads.
+For distributed setup there is an high initial latency due to communication overheads.
+After that it comes to a constant value.
 
@@
 Finally \textit{TopProfitProcessor} aggregates events and writes to a file.
-Similar to earlier experiment,  we deployed this process graph to our system and measured the execution time and the average message latency with the data set given for both single JVM setup and distributed setup.
+
+Similar to earlier experiment,  we deployed this process graph to our system and measured the throughput and the average message latency increasing the number of  \textit{ProfitCalculaor} instances for both single JVM setup and distributed setup to examine the scalability of our solution.
 
@@
         \centering
-        \includegraphics[width=3.0in]{time_profit.png}
-        \caption{Execution times to evaluate top ten profitable cells.}
-        \label{time_profit}
+        \includegraphics[width=3.0in]{throughput_profit.png}
+        \caption{Throughput variation with the number of \textit{ProfitCalculator} instances.}
+        \label{throughput_profit}
 \end{figure}
 
-Figure \ref{time_profit} shows the execution times for both setups.
-In both setups, there is a higher execution time at one instance due to lack of parallelism.
-Then for distributed set up execution times get reduced with addition of new resources and for single JVM setup it remains a constant.
-This shows our solution scales up in a distributed setup.
-We can increase the system throughput by adding new nodes until \textit{ProfitEventEmitter} becomes the bottleneck.
+Figure \ref{throughput_profit} shows the throughput variation for both setups.
+For single JVM set up initially throughput increases, due to increment of parallelism and comes to a steady state.
+For distributed setup  throughput increases linearly, although it has an initial lesser throughput compared to single JVM setup due to message serialization, message deserialization and tcp overheads.
 
@@
         \includegraphics[width=3.0in]{latency_profit.png}
-        \caption{Latencies to evaluate top ten profitable cells.}
+        \caption{Mean latency variation with the number of \textit{ProfitCalculator} instances.}
         \label{latency_profit}
@@
 
-Figure \ref{latency_profit} shows the latencies.
+Figure \ref{latency_profit} shows the latency variation for both setups.
 For single JVM, latencies slightly increased with the addition of new instances as expected.
-But for distributed setups there is a higher latencies compared to others.
-This latency is due to longer queues at the \textit{TopProfitProcessor}.
+But for distributed setups latency linearly increases with the addition of new nodes after 6 \textit{ProfitCalculator} instances.
+This latency increment is due to longer queues at the \textit{TopProfitProcessor}.
 In our message ordering process, we can send a message to \textit{TopProfitProcessor} only if all queues are not empty.

### 2015-03-26 paper/experiments.tex
fixing meanings

@@
 We conducted several experiments to study the scalability of our solution.
-Firstly we study the throughput and mean message latency variation with the number of \textit{QueryEvaluators} for both single JVM and distributed setups.
+Firstly we study the throughput and mean message latency variation with the number of \textit{QueryEvaluators} for both single JVM and distributed setups to find the scalability of the parallel evaluation.
 In an ideal system, we should observe linear throughput increment with a constant message latency.
@@
 Secondly, there is only one \textit{RouteEventEmitter} and the speed at which it can read and send messages becomes the bottleneck.
-Initial throughput for distributed setup is less, due to message serialization, message deserialization and tcp overheads.
+Initial throughput for distributed setup is less, due to message serialization, message deserialization and TCP overheads.
 
@@
         \centering
-        \includegraphics[width=3.0in]{profictgraph.png}
+        \includegraphics[width=3.0in]{profitgraph.png}
         \caption{Process graph to evaluate top ten profitable cells.}
@@
 For single JVM set up initially throughput increases, due to increment of parallelism and comes to a steady state.
-For distributed setup  throughput increases linearly, although it has an initial lesser throughput compared to single JVM setup due to message serialization, message deserialization and tcp overheads.
+For distributed setup  throughput increases linearly, although it has an initial lesser throughput compared to single JVM setup due to message serialization, message deserialization and TCP overheads.
 

### 2015-03-26 paper/query_evaluation.tex
fixing meanings

@@
 
-In this section, we explain how to evaluate queries with the given dataset.
+In this section, we explain how to evaluate queries with the given data set.
 For this query evaluation we can use a distributed event stream processing system with the process graph as shown in Figure \ref{twonode_graph} for both queries.
@@
 Since this is not the desired result, we can add another process to aggregate these top result events and generate global route change events.
-Figure \ref{parallel} shows the possible execution patterns with the new process.
+Figure \ref{parallel} shows, possible parallel evaluation techniques with the new process.
+Multiple \textit{QueryEvaluator} instances can be executed in a single JVM using multiple threads or in a distributed setup with many JVMs using TCP to communicate.
 

### 2015-03-26 paper/query_implementation.tex
fixing meanings

@@
 
+\subsection{Finding Runtime Median}
+In general, we can find runtime median using two heaps.
+One to keep values greater than the current median and one to keep values less than the current median.
+An implement of this algorithm using two \textit{java.util.PriorityQueue}s have an O(log(n)) time complexity for add operation and O(n) time complexity for remove operation (\texttt{delete} operation of \textit{java.util.PriorityQueue} takes O(n) time).
+We overcame this issue using \textit{DynamicHeap} to keep values.
+In our solution, we use value as the key to insert and delete from the \textit{DynamicHeap} and keep value with its count (same value can be repeated) in the \textit{DynamicHeap}.
+In worst case it has O(log(n)) time complexity for both add and remove operations and in best case (i.e value already present and that does not cause median change) it has O(1) time complexity.
+
 \subsection{Frequent Route Query}
@@
 Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to process an event.
-This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an Object called \textit{RouteCount}.
+This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an object called \textit{RouteCount}.
 
@@
 	\item The profitability is calculated using the median profit and the number of empty taxis.
-In order to calculate median with O(log(n)) time complexity, we use two heaps (\textit{java.util.PriorityQueue}) .
-One is used to keep the values lesser than the median value (max heap) and other is to keep higher values (min heap).
-However the \texttt{delete} operation of \textit{java.util.PriorityQueue} takes O(n) time, since it traverse along the array to find the index of the object.
-We implemented a similar algorithm using our \textit{DynamicHeap} to keep higher and lower values than median with value count.
-In worst case it takes O(log(n)) time to add or remove a value, in best case (i.e if the value is already present and that does not cause median to change) it takes O(1) time.
-But that did not improved the performance.
+In order to calculate median, we use \textit{java.util.PriorityQueue} based approach mentioned in (Section 2.4).
+We choose this option over our implementation, because this approach performs better with smaller window sizes.
 First, these heap sizes do not grow over 145 and there are same values repeating in the list increasing the probability of finding a value without traversing the whole list.
-Secondly, our implementation create more objects compared to just keeping double values though the time complexity is O(log(n)).
+Secondly, our implementation create more objects per operation compared to just keeping double values though the time complexity is O(log(n)).
+In section 4.3 we show that our approach works better for larger window sizes.
 	\item In order to keep track of empty taxis, we need to process both pickup cell and dropoff cell details for each message.

### 2015-03-26 src/assembly/gc.sh
creating a distribution

@@ -1,0 +1,3 @@
+#! /bin/bash
+
+java -cp lib/grand_challenge-1.0.jar:lib/stream-1.0.jar edu.colostate.cs.gc.GCEventEmitter $1 2 3

### 2015-03-26 src/assembly/profit.sh

@@ -1,0 +1,3 @@
+#! /bin/bash
+
+java -cp lib/grand_challenge-1.0.jar:lib/stream-1.0.jar edu.colostate.cs.gc.profit.ProfitEventEmitter $1 4 1

### 2015-03-26 src/assembly/route.sh

@@ -1,0 +1,3 @@
+#! /bin/bash
+
+java -cp lib/grand_challenge-1.0.jar:lib/stream-1.0.jar edu.colostate.cs.gc.route.RouteEventEmitter $1 2 

### 2015-03-26 src/main/java/edu/colostate/cs/gc/GCEventEmitter.java

@@ -146,9 +146,9 @@ public class GCEventEmitter {
             }
 
             long totalTime = System.currentTimeMillis() - currentTime;
-            System.out.println("Display statistics ...");
-            System.out.println("Total time (ms) " + totalTime);
-            System.out.println("Total time (Min) " + totalTime * 1.0 / 60000);
+            System.out.println("Execution time (both Queries) ...");
+            System.out.println("Execution time (ms) " + totalTime);
+            System.out.println("Execution time (Min) " + totalTime * 1.0 / 60000);
 
             bufferedReader.close();
 
@@ -219,7 +219,9 @@ public class GCEventEmitter {
         new GCEventEmitter().loadData(args[0], routeBuffers, profitBuffers, barrier, latch);
 
         topRouteProcessor.close();
+        System.out.println("Route query delay ...");
         Util.displayStatistics("top_routes.txt");
+        System.out.println("Profit cell query delay ...");
         topProfitProcessor.close();
         Util.displayStatistics("top_profit_cells.txt");
 

### 2015-03-26 src/main/java/edu/colostate/cs/gc/profit/ProfitCalculator.java

@@ -49,6 +49,7 @@ public class ProfitCalculator extends TripProcessor {
     public void initialise(Container container, Map<String, String> parameters) {
         super.initialise(container, parameters);
         this.numOfProcessors = Integer.parseInt(parameters.get("processors"));
+        this.windowSize = Constants.SMALL_WINDOW_SIZE;
     }
 
     public void processEvent(TripEvent event) {

### 2015-03-26 src/main/java/edu/colostate/cs/gc/profit/ProfitEventEmitter.java

@@ -143,8 +143,8 @@ public class ProfitEventEmitter implements Adaptor {
 
             long totalTime = System.currentTimeMillis() - currentTime;
             System.out.println("Display statistics ...");
-            System.out.println("Total time (ms) " + totalTime);
-            System.out.println("Total time (Min) " + totalTime * 1.0 / 60000);
+            System.out.println("Execution time (ms) " + totalTime);
+            System.out.println("Execution time (Min) " + totalTime * 1.0 / 60000);
             System.out.println("Throughput (Msg/s) " + numberOfEvents * 1000.0 / totalTime);
 
             bufferedReader.close();

### 2015-03-26 src/main/java/edu/colostate/cs/gc/route/RouteEventEmitter.java

@@ -97,8 +97,8 @@ public class RouteEventEmitter implements Adaptor {
 
             long totalTime = System.currentTimeMillis() - currentTime;
             System.out.println("Display statistics ...");
-            System.out.println("Total time (ms) " + totalTime);
-            System.out.println("Total time (Min) " + totalTime * 1.0 / 60000 );
+            System.out.println("Execution time (ms) " + totalTime);
+            System.out.println("Execution time (Min) " + totalTime * 1.0 / 60000 );
             System.out.println("Throughput (Msg/s) " + numberOfEvents * 1000.0 / totalTime);
 
             bufferedReader.close();

### 2015-03-26 src/main/java/edu/colostate/cs/gc/util/Util.java

@@ -74,8 +74,7 @@ public class Util {
                      break;
                 }
             }
-            System.out.println("Total lines " + totalLines);
-            System.out.println("Average delay " + totalDelay * 1.0 / totalLines);
+            System.out.println("Mean delay " + totalDelay * 1.0 / totalLines);
         } catch (FileNotFoundException e) {
             e.printStackTrace();  //To change body of catch statement use File | Settings | File Templates.
         } catch (IOException e) {

### 2015-03-28 paper/introduction.tex
add introduction

\section{Introduction}
Stream processing systems are particularly suitable for processing tuples that may be reported by devices, computing processes, or humans. These streams are processed to perform knowledge extraction based on the information encoded within individual tuples. Given the rate at which these tuples are generated, inability of the processing to keep pace with the data generation rates will lead to buffer overflows, dropped packets, and incorrect results. Often the data from these streams must be processed in real time and the knowledge based updated as well.

In this paper, we provide our solution to the DEBS Grand Challenge problem of processing New York taxi trip data and to perform knowledge extraction. We consider this problem in the context of stream processing where the tuples are passed through stream processing stages and the knowledge base updated in real time.
\subsection{Research Challenges}
Processing trip data streams and continually updating the knowledge base includes the following challenges:
\begin{enumerate}
	\item \textit{Data is continually arriving.} The processing of the data streams must keep pace with the data generation rates.
	\item \textit{Regardless of the data arrival rates, at any given instant the knowledge base must be accurate and current:} Even as tuples continually arrive, the information desired by the system (the top 10 values) must reflect the current operating conditions. It is possible that as a new trip information arrives, the list of top 10 routes may change. This update must occur in real time as soon as the tuple that triggered this change is processed.
	\item \textit{Updates to data structures must be accurate and keep pace with data arrival rates.} The time and space complexity of the data structures used to manage the knowledge base must be taken into account. Specifically, the data structure that we use must not preclude real time processing.
\end{enumerate}

\subsection{Research Questions}
We consider several research questions in our solution to the DEBS Grand Challenge problem. These include:
\begin{enumerate}
	\item \textit{How can we support real-time analytics and processing over these streams?}  Processing at this scale and in real-time crosscuts algorithmic and system design. The framework must be able to scale with increase in data volumes and arrival rates.
	\item \textit{How can we extract and update knowledge base in real time?}  Since the data streams are continually arriving the space and time complexity of the data structures become important. The data structures must be compact to conserve memory and the time-complexity of the operations to update the structures must be low.
	\item \textit{How can we ensure high-throughput processing to support processing a large number of packets per second?} This could involve leveraging multiple instances of various processing stages. 
\end{enumerate}

\subsection{Contributions}
We observe following key contributions, which can be used in solving future DEBS challenges as well.
\begin{enumerate}
	\item We have developed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log(n)) time complexity. 
	\item Using our data structures we have developed a new algorithm to find the running median with O(log(n)) time complexity. One of the key challenges of DEBS 2014 \cite{jerzak2014debs}  was to find the median of a large time window. We believe our solution is useful in such situations.
	\item We have developed a scalable parallel evaluation technique to process queries efficiently. 
\end{enumerate}

\subsection{Paper Organization}
The remainder of this paper is organized as follows. Section 2 provides an in-depth description about the internal data structures and the algorithms used to process the events. Section 3 describes our scalable design to process events and associated problems. Section 4 proves the soundness of our solution with experimental results. Section 5 discuss the related work including some last year solutions. Section 6 summarizes key concepts and discuss their further applications.

### 2015-03-28 paper/introduction.tex
updating the introduction

@@
 \section{Introduction}
-Stream processing systems are particularly suitable for processing tuples that may be reported by devices, computing processes, or humans.
-These streams are processed to perform knowledge extraction based on the information encoded within individual tuples.
-Given the rate at which these tuples are generated, inability of the processing to keep pace with the data generation rates will lead to buffer overflows, dropped packets, and incorrect results.
-Often the data from these streams must be processed in real time and the knowledge based updated as well.
 
-In this paper, we provide our solution to the DEBS Grand Challenge problem of processing New York taxi trip data and to perform knowledge extraction.
-We consider this problem in the context of stream processing where the tuples are passed through stream processing stages and the knowledge base updated in real time.
-\subsection{Research Challenges}
-Processing trip data streams and continually updating the knowledge base includes the following challenges:
+The Grand Challenge 2015 problem focuses on processing New York city taxi trip data for year 2013.
+The area considered is a square of 150m x  150m.
+The  square is divided into smaller units called cells and a taxi trip is considered as moving from one cell (pickup cell) to another cell (dropoff cell).
+Therefore a route is uniquely identified by the pickup cell and the dropoff cell.
+The data set consists of 173 million events and solutions are expected to replay these data and generate events that meet criteria of two queries defined.
+
+The first query is to find the events which change the top ten routes within last 30 minutes and emit the new route set details along the original event.
+The given data set contains 899010  routes and the 30 minute time window for this query accounts around 10000 events assuming an equal event distribution.
+This creates the possibility of processing several thousand route counts specially within peak ours.
+In other words we need to evaluate several thousand route counts with each event and find 10 top most routes.
+Similar analysis can be made with profitable cells query as well.
+In addition to that it requires to  calculate the median fare for each cell.
+
+\subsection{Research Challengers}
+
+If we consider an application to process this 173 million data set within an year, it would be a trivial task and any naive implementation satisfy that requirement.
+However if we consider an application which process these data within minutes, we observe following research challengers.
 \begin{enumerate}
-	\item \textit{Data is continually arriving.} The processing of the data streams must keep pace with the data generation rates.
-	\item \textit{Regardless of the data arrival rates, at any given instant the knowledge base must be accurate and current:} Even as tuples continually arrive, the information desired by the system (the top 10 values) must reflect the current operating conditions.
-It is possible that as a new trip information arrives, the list of top 10 routes may change.
-This update must occur in real time as soon as the tuple that triggered this change is processed.
-	\item \textit{Updates to data structures must be accurate and keep pace with data arrival rates.} The time and space complexity of the data structures used to manage the knowledge base must be taken into account.
-Specifically, the data structure that we use must not preclude real time processing.
+	\item \textit{Handling thousands of values in real time :}  An algorithm with O(n) time complexity will not scale for thousands.
+Therefore is it possible to come up with an algorithm of O(log(n)) time complexity?
+	\item \textit{Parallel evaluation :}  This application is not inherently support parallel evaluation.
+Partitioning route sets and cell sets triggers events for sub sets, but application requires the events with whole sets.
+Therefore is it possible to partition data and still generate required results to support parallelism?
 \end{enumerate}
 
-\subsection{Research Questions}
-We consider several research questions in our solution to the DEBS Grand Challenge problem. These include:
-\begin{enumerate}
-	\item \textit{How can we support real-time analytics and processing over these streams?}  Processing at this scale and in real-time crosscuts algorithmic and system design.
-The framework must be able to scale with increase in data volumes and arrival rates.
-	\item \textit{How can we extract and update knowledge base in real time?}  Since the data streams are continually arriving the space and time complexity of the data structures become important.
-The data structures must be compact to conserve memory and the time-complexity of the operations to update the structures must be low.
-	\item \textit{How can we ensure high-throughput processing to support processing a large number of packets per second?} This could involve leveraging multiple instances of various processing stages. 
-\end{enumerate}
+\subsection{Methodology}
+
+We started our work seeking answers for above challengers.
+First we evaluated doubly linked list and heap based approaches to build a data structure with O(log(n)) time complexity.
+After several evaluations, we developed a new data structure called \textit{NodeList} which utilizes both concepts.
+In order to find the running median with O(log(n)) time complexity, we experimented with an algorithm which uses priority queue and the heap data structure we developed.
+We decided to use former since it gave better performance with smaller window sizes.
+Then we focused on parallelizing the application and came up with an strategy to aggregate the sub set results to find the final results.
+As the last step, we conducted several experiments to evaluate our solutions.
+We repeated this process identifying possible issues and fixing them.
 

### 2015-03-30 paper/conclustion.tex
fixing some issues

\section{Conclusion}

We propose a scalable solution for DEBS grand challenge problem 2015. Our queries execute with O(log(n)) time complexity and hence scales with window size. We propose a scalable evaluation technique to execute queries in a distributed setup. Since \textit{EventEmitter} partitions the events using the  hash code value of the key (\textit{Route} or \textit{Cell}), our solution can be used to process events with multiple \textit{EventEmitters} (multiple event streams) as well. Different levels of \textit{ResultAggregators} can be used to scales up aggregation process if required. We propose an algorithm to find the runtime median with O(log(n)) time complexity. This algorithm specially suites for finding median with repeating values. The time windows used in our solution has an O(1) time complexity but O(n) space complexity. This creates opportunity to research further on efficient memory management techniques for large window sizes.

### 2015-03-30 paper/experiments.tex
fixing some issues

@@
 Firstly we study the throughput and mean message latency variation with the number of \textit{QueryEvaluators} for both single JVM and distributed setups to find the scalability of the parallel evaluation.
+We measured the end to end latency.
+At the \textit{EventEmitter}, we set the time stamp for all events once the record is read.
+At the \textit{QueryEvaluator} we copy the original time stamp, if that event causes a top ten value change event.
+Total delay calculated at the \textit{ResultAggregator} just before writing the event.
 In an ideal system, we should observe linear throughput increment with a constant message latency.
 However there are some exceptions and we have provided our explanations on them.
-Finally we study the throughput variation with the window size to study the suitability of our solution for higher window sizes.
+Finally we studied the throughput variation with the window size to study the suitability of our solution for higher window sizes.
 All these experiments were performed on a LAN with a network bandwidth of 1Gbps and all machines were Intel(R) Xeon(R) 2.4GHz, 4 core duo machines with 16 GB of memory.

### 2015-03-30 paper/introduction.tex
fixing some issues

@@
 The given data set contains 899010  routes and the 30 minute time window for this query accounts around 10000 events assuming an equal event distribution.
-This creates the possibility of processing several thousand route counts specially within peak ours.
+This creates the possibility of processing several thousand route counts specially within peak hours.
 In other words we need to evaluate several thousand route counts with each event and find 10 top most routes.
@@
 If we consider an application to process this 173 million data set within an year, it would be a trivial task and any naive implementation satisfy that requirement.
-However if we consider an application which process these data within minutes, we observe following research challengers.
+However if we consider an application which process these data within minutes with minimal latency to process each event, we observe following research challengers.
 \begin{enumerate}
@@
 We decided to use former since it gave better performance with smaller window sizes.
-Then we focused on parallelizing the application and came up with an strategy to aggregate the sub set results to find the final results.
-As the last step, we conducted several experiments to evaluate our solutions.
+Then we focused on parallelizing the application and came up with a strategy to aggregate the sub set results to find the final results.
+As the last step, we conducted several experiments to evaluate the scalability of our solution.
 We repeated this process identifying possible issues and fixing them.
@@
 Section 2 provides an in-depth description about the internal data structures and the algorithms used to process the events.
-Section 3 describes our scalable design to process events and associated problems.
-Section 4 proves the soundness of our solution with experimental results.
+Section 3 describes our scalable design to evaluate queries and associated problems.
+Section 4 proves the scalability of our solution with experimental results.
 Section 5 discuss the related work including some last year solutions.

### 2015-03-30 paper/paper.tex
fixing meanings
fixing some issues

@@
 
-\title{DBES Grand Challenge : High throughput, scalable solution for processing taxi trip data}
+\title{A High-throughput, Scalable Solution for the New York Taxi Trips DEBS Grand Challenge Problem}
 %
@@
 Firstly, such processing demands efficient data structures and algorithms to minimize the processing overhead.
-Secondly, the system should partition data to process them in parallel to make it scalable.
+Secondly, the system should partition data to evaluate them in parallel to make it scalable.
 
-In this paper, we present a set of data structures which can be used to evaluate the aforementioned queries with O(log(n)) time complexity and a data partitioning technique to process events in parallel.
-Then we evaluate our solution within signal node as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN.
+In this paper, we present a set of data structures which can be used to evaluate the aforementioned queries with O(log(n)) time complexity and a data partitioning technique to evaluate them in parallel.
+Then we evaluate our solution in a signal machine as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN.
 We could process frequent routes query with 173 million data set within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
@@
 
-\section{Introduction}
-%introduction
-
+\input{introduction}
 \input{query_implementation}
+\clearpage
 \input{query_evaluation}
@@
 \input{related_work}
+\input{conclustion}
 %

### 2015-03-30 paper/query_evaluation.tex
fixing some issues

@@
 
-We can sequentially execute events either in a single Java Virtual Machine (JVM) or different JVMs (either in one machine or different machines) using TCP sockets to communicate among process instances.
+We can sequentially execute events either in a single Java Virtual Machine (JVM) or in different JVMs (either in one machine or different machines) using TCP sockets to communicate among process instances.
 Figure \ref{sequential} shows possible executions.
@@
 For an example, lets assume route count for route R1 is 10 and R1 is included in a top ten route event generated by a \textit{QueryEvaluator}.
-Then lets assume route count for R1 has been reduced to 7 (due to expiration events), when the next event is generated and hence R1 is not included in the next top ten route event.
+Then lets assume when the next event is generated, the route count for R1 has been reduced to 7 (due to expiration events) and hence R1 is not included in the next top ten route event.
 However since R1 value still is in the \textit{ResultAggregator} that may produce an event with this value.

### 2015-03-30 paper/query_implementation.tex
fixing some issues

@@
 First is to evaluate top ten most frequent routes within the last 30 minutes and second is to evaluate top ten profitable areas within the last 15 minutes.
-We use a composite object to manage details for each route and fare details to each cell to compare values among routes and cells.
+We can use composite objects to manage details for each route and cell and compare them to find top 10 values.
 However, each time window can have thousands of route and cell details.
@@
 In order to support all these functions with O(log(n)) time complexity we have developed a data structure called \textit{NodeList}.
-The remaining of this section describes how we have achieve this time complexity with the \textit{NodeList} data structure and how we have implemented the queries using that data structure.
+The remaining of this section describes how we have achieved this time complexity with the \textit{NodeList} data structure and how we have implemented the queries using that data structure.
 
@@
 For an example, the route query can use a \textit{RouteCount} object to keep route count while profitable cells query can use \textit{CellProfit} object to keep track of fare details.
-This interface contains a method called compare to define the order of the values.
+This interface contains a method called \texttt{compare} to define the order of the values.
 The \textit{NodeList} data structure provides the notion of a sorted values starting from position 0 (highest value) to its users.
@@
 
-\subsection{Finding Runtime Median}
-In general, we can find runtime median using two heaps.
+\subsection{Finding Runing Median}
+In general, we can find runing median using two heaps.
 One to keep values greater than the current median and one to keep values less than the current median.
-An implement of this algorithm using two \textit{java.util.PriorityQueue}s have an O(log(n)) time complexity for add operation and O(n) time complexity for remove operation (\texttt{delete} operation of \textit{java.util.PriorityQueue} takes O(n) time).
+An implement of this algorithm using two \textit{java.util.PriorityQueue}s have an O(log(n)) time complexity for add operation and O(n) time complexity for remove operation (\texttt{remove} operation of \textit{java.util.PriorityQueue} takes O(n) time).
 We overcame this issue using \textit{DynamicHeap} to keep values.
@@
 Since queue operations are supported in O(1) time, query evaluations happen in O(log(n)) time.
-Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to process an event.
+Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to evaluate the query an event.
 This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an object called \textit{RouteCount}.
@@
 
-Processing an event can cause add, remove or update values of the existing time window objects.
+Each query evaluation can cause add, remove or update values of the existing time window objects.
 Therefore query processing mainly involves performing the above operations with the underlying \textit{NodeList} data structure and comparing top ten routes before and after.
@@
 First, these heap sizes do not grow over 145 and there are same values repeating in the list increasing the probability of finding a value without traversing the whole list.
-Secondly, our implementation create more objects per operation compared to just keeping double values though the time complexity is O(log(n)).
+Secondly, our implementation creates more objects per operation compared to just keeping double values though the time complexity is O(log(n)).
 In section 4.3 we show that our approach works better for larger window sizes.

### 2015-03-30 paper/related_work.tex
fixing meanings
fixing some issues

@@
 All these systems are based on the Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data.
-Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence dynamic membership management.
-Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets RDDs \cite{zaharia2012resilient} and Muppet \cite{lam2012muppet} provides state persistence using a \textless key, value\textgreater store based on Cassandra \cite{lakshman2010cassandra}.
+Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence and dynamic membership management.
 

### 2015-04-28 paper/conclustion.tex
commiting last changes

@@
-\section{Conclusion}
+\section{Conclusions and Future Work}
 
 We propose a scalable solution for DEBS grand challenge problem 2015.
-Our queries execute with O(log(n)) time complexity and hence scales with window size.
+Our queries execute with O(log n) time complexity and hence scale with window size.
 We propose a scalable evaluation technique to execute queries in a distributed setup.
 Since \textit{EventEmitter} partitions the events using the  hash code value of the key (\textit{Route} or \textit{Cell}), our solution can be used to process events with multiple \textit{EventEmitters} (multiple event streams) as well.
-Different levels of \textit{ResultAggregators} can be used to scales up aggregation process if required.
-We propose an algorithm to find the runtime median with O(log(n)) time complexity.
-This algorithm specially suites for finding median with repeating values.
-The time windows used in our solution has an O(1) time complexity but O(n) space complexity.
-This creates opportunity to research further on efficient memory management techniques for large window sizes.
+Different levels of \textit{ResultAggregators} can be used to scale up aggregation process if required.
+We propose an algorithm to find the runtime median with O(log n) time complexity.
+This algorithm is especially suited for finding median with repeating values.
+The time windows used in our solution have an O(1) time complexity, but O(n) space complexity.
+This creates opportunities for future research on efficient memory management techniques for large window sizes.

### 2015-04-28 paper/experiments.tex
commiting last changes

@@
 We conducted several experiments to study the scalability of our solution.
-Firstly we study the throughput and mean message latency variation with the number of \textit{QueryEvaluators} for both single JVM and distributed setups to find the scalability of the parallel evaluation.
+We profiled the throughput and mean message latency variation with the number of \textit{QueryEvaluators} for both single JVM and distributed setups to find the scalability of the parallel evaluation.
 We measured the end to end latency.
@@
 At the \textit{QueryEvaluator} we copy the original time stamp, if that event causes a top ten value change event.
-Total delay calculated at the \textit{ResultAggregator} just before writing the event.
+Total delay is calculated at the \textit{ResultAggregator} just before writing the event.
 In an ideal system, we should observe linear throughput increment with a constant message latency.
@@
 
-\subsection{Frequent routes query}
+\subsection{Frequent Routes Query}
 
@@
 
-We deployed this process graph to a distributed stream processing system we have developed, and  measured the throughput and the mean message latency varying the number of \textit{RouteProcessors} with the data set given (173 million records).
+We deployed this process graph to a distributed stream processing system we have developed, and  measured the throughput and the mean message latency varying the number of \textit{RouteProcessors} with the dataset given (173 million records).
 In the local set up, we use a new thread to execute a new instance and in the distributed setup we add a new node to execute the new instance.
@@
 For single JVM setup latency get reduced at 2 and slightly increases after that due to context switching overhead with higher number of threads.
-For distributed setup there is an high initial latency due to communication overheads.
+For distributed setup there is a high initial latency due to communication overheads.
 After that it comes to a constant value.
 
-\subsection{Profitable cells query}
+\subsection{Profitable Cells Query}
 
@@
 In this way cells are partitioned into different \textit{ProfitCalculator} instances and each instance gets all the events relevant to its' cell set.
-We use a set of threads (equals to the number of Profict Calculators) to emit data as in the previous query.
+We use a set of threads (equals to the number of \textit{ProfitCalculators}) to emit data as in the previous query.
 Finally \textit{TopProfitProcessor} aggregates events and writes to a file.
@@
 The grand challenge problem uses 15 minute window size for profitable cells query.
-However in a real world problems window sizes may required.
+However in real world problems higher window sizes may required.
 Therefore finally we evaluated scalability of our solution with the window size.
@@
         \includegraphics[width=3.0in]{window.png}
-        \caption{Throughput variation with the window size.}
+        \caption{Throughput variation with window size.}
         \label{window}
@@
 For both solutions, throughput does not drop by half, even though the window size increases 64 times.
-And for large windows \textit{DynamicHeap} performs better than the \textit{PriorityQueue} based one.
+For larger windows \textit{DynamicHeap} performs better than the \textit{PriorityQueue} based one.
 This is due to two reasons.
-Firstly, all operations of \textit{DynamicHeap} are O(log(n)).
-Secondly, out of 173 million records there are only around 18000 unique fares.
-Therefore \textit{DyamicHeap} based algorithm does not consume memory beyond that.
-However , since \textit{PriortyQueue} based implementation keeps all values in memory it consumes more memory for large windows.
+First, all operations of \textit{DynamicHeap} are O(log n).
+Second, out of the 173 million records there are only around 18000 unique fares.
+Therefore the \textit{DyamicHeap} based algorithm does not consume memory beyond that.
+However, since \textit{PriortyQueue} based implementation keeps all values in memory, it consumes more memory for larger windows.

### 2015-04-28 paper/introduction.tex
commiting last changes

@@
 
-The Grand Challenge 2015 problem focuses on processing New York city taxi trip data for year 2013.
-The area considered is a square of 150m x  150m.
+The Grand Challenge 2015 problem focuses on processing New York city taxi trip data for the year 2013.
+The area considered is a square of 150km x  150km.
 The  square is divided into smaller units called cells and a taxi trip is considered as moving from one cell (pickup cell) to another cell (dropoff cell).
-Therefore a route is uniquely identified by the pickup cell and the dropoff cell.
-The data set consists of 173 million events and solutions are expected to replay these data and generate events that meet criteria of two queries defined.
+A route is uniquely identified by the pickup cell and the dropoff cell.
+The dataset consists of 173 million events and solutions are expected to replay these data and generate events that meet the criteria of two defined queries.
 
-The first query is to find the events which change the top ten routes within last 30 minutes and emit the new route set details along the original event.
-The given data set contains 899010  routes and the 30 minute time window for this query accounts around 10000 events assuming an equal event distribution.
-This creates the possibility of processing several thousand route counts specially within peak hours.
-In other words we need to evaluate several thousand route counts with each event and find 10 top most routes.
-Similar analysis can be made with profitable cells query as well.
+The first query is to find the events that change the top ten routes within last 30 minutes and emit the new route set details along the original event.
+The given dataset contains 899010 routes and the 30 minute time window for this query accounts around 10000 events assuming an equal event distribution.
+This creates the possibility of processing several thousand route counts especially during peak hours.
+In other words we need to evaluate several thousand route counts with each event and find 10 top routes.
+A similar analysis can be made for the profitable cells query as well.
 In addition to that it requires to calculate the median fare for each cell.
 
-\subsection{Research Challengers}
+\subsection{Research Challenges}
 
-If we consider an application to process this 173 million data set within an year, it would be a trivial task and any naive implementation satisfy that requirement.
-However if we consider an application which process these data within minutes with minimal latency to process each event, we observe following research challengers.
+If we consider an application that must process the 173 million trips dataset within an year, it would be a trivial task and several simple implementations would satisfy that requirement.
+However if we consider an application that must process these data within minutes with the added constraint of minimal latency to process each event, we observe following research challenges.
 \begin{enumerate}
-	\item \textit{Handling thousands of values in real time :}  An algorithm with O(n) time complexity will not scale for thousands.
-Therefore is it possible to come up with an algorithm of O(log(n)) time complexity?
-	\item \textit{Parallel evaluation :}  This application is not inherently support parallel evaluation.
-Partitioning route sets and cell sets triggers events for sub sets, but application requires the events with whole sets.
+	\item \textit{Handling thousands of values in real time :}  An algorithm with O(n) time complexity will not scale for thousands. How can we design an algorithm with O(log n) time complexity?
+	\item \textit{Parallel evaluation :}  This application does not inherently support parallel evaluation.
+Partitioning route sets and cell sets triggers events for subsets, but application requires the events with whole sets.
 Therefore is it possible to partition data and still generate required results to support parallelism?
@@
 
-We started our work seeking answers for above challengers.
-First we evaluated doubly linked list and heap based approaches to build a data structure with O(log(n)) time complexity.
-After several evaluations, we developed a new data structure called \textit{NodeList} which utilizes both concepts.
-In order to find the running median with O(log(n)) time complexity, we experimented with an algorithm which uses priority queue and the heap data structure we developed.
+We started our work seeking answers for the above challenges.
+First we evaluated doubly linked list and heap based approaches to build a data structure with O(log n) time complexity.
+After several evaluations, we developed a new data structure called \textit{NodeList} that utilizes both these concepts.
+In order to find the running median with O(log n) time complexity, we experimented with an algorithm which uses priority queue and the heap data structure we developed.
 We decided to use former since it gave better performance with smaller window sizes.
-Then we focused on parallelizing the application and came up with a strategy to aggregate the sub set results to find the final results.
-As the last step, we conducted several experiments to evaluate the scalability of our solution.
-We repeated this process identifying possible issues and fixing them.
+We then focused on parallelizing the application and designed a strategy to aggregate the subset results to find the final results.
+We also conducted several experiments to evaluate the scalability of our solution.
 
@@
 \begin{enumerate}
-	\item We have developed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log(n)) time complexity.
-	\item Using our data structures we have developed a new algorithm to find the running median with O(log(n)) time complexity.
+	\item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity.
+	\item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity.
 One of the key challenges of DEBS 2014 \cite{jerzak2014debs}  was to find the median of a large time window.
@@
 Section 4 proves the scalability of our solution with experimental results.
-Section 5 discuss the related work including some last year solutions.
-Section 6 summarizes key concepts and discuss their further applications.
+Section 5 discusses the related work including some last year solutions.
+Section 6 summarizes key concepts and discusses their further applications.

### 2015-04-28 paper/paper.tex
commiting last changes

@@
 \begin{abstract}
-Processing complex queries on unbounded event streams in real time, is a challenge for many organizational data processing systems.
-These systems are expected to process data with reduced latency to generate real time events, and high throughput to minimize the required hardware.
+Processing complex queries on unbounded event streams in real time, is a challenge for many data processing systems.
+These systems are expected to process data with reduced latency to generate real time-events, and at high throughput to minimize the required hardware.
 In this regard, DEBS Grand Challenge 2015 focuses on evaluating two queries (frequent routes and  profitable cells) in real time with low latency and high throughput.
@@
 
-In this paper, we present a set of data structures which can be used to evaluate the aforementioned queries with O(log(n)) time complexity and a data partitioning technique to evaluate them in parallel.
-Then we evaluate our solution in a signal machine as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN.
-We could process frequent routes query with 173 million data set within 5 minutes with less than 4 millisecond latency and profitable cells query with same data set within 11 minutes with less than 5 millisecond latency.
+In this paper, we present a set of data structures that we designed to evaluate the aforementioned queries with O(log n) time complexity and a data partitioning technique to evaluate them in parallel.
+We then evaluate our solution on a signal machine as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN.
+We are able to process the frequent routes query with the 173 million trips dataset within 5 minutes with less than 4 millisecond latency and the profitable cells query with same dataset within 11 minutes with less than 5 millisecond latency.
 \end{abstract}

### 2015-04-28 paper/query_evaluation.tex
commiting last changes

@@
 
-\subsection{Sequential evaluation}
+\subsection{Sequential Evaluation}
 
@@
 
-First both processors can be executed in same JVM with same thread.
+First both processes can be executed in same JVM with same thread.
 In that case, both processes execute sequentially.
@@
 
-\subsection{Parallel evaluation}
+\subsection{Parallel Evaluation}
 The general way to parallelize event processing, is to partition the events according to a key and run multiple instances to process data.
@@
 
-Algorithm \ref{aggregate_algorithm} shows the algorithm for ResultAggregator with above fixes.
-First it removes the expired keys.
+Algorithm \ref{aggregate_algorithm} depicts the algorithm for \textit{ResultAggregator} with the above fixes.
+The algorithm first removes the expired keys.
 Then it updates the existing values or adds new values and moves the position in the list accordingly.

### 2015-04-28 paper/query_implementation.tex
commiting last changes

@@
 \section{Query Implementation}
-In this section, we elaborate how we have implemented the query logic to achieve O(log(n)) time complexity.
+In this section, we elaborate how we have implemented the query logic to achieve O(log n) time complexity.
 The 2015 Grand Challenge problem has two queries.
@@
 However, each time window can have thousands of route and cell details.
-Therefore, an efficient data structure is required to add, remove, update and retrieve top ten values from a set of objects for a given time window efficiently.
-Further, since each event is received with route and cell details, the aforementioned operations should be performed using route and cell details as the key.
-In order to support all these functions with O(log(n)) time complexity we have developed a data structure called \textit{NodeList}.
-The remaining of this section describes how we have achieved this time complexity with the \textit{NodeList} data structure and how we have implemented the queries using that data structure.
+Therefore, an efficient data structure is needed to add, remove, update and retrieve top ten values from a set of objects for a given time window efficiently.
+Since each event is received with route and cell details, the aforementioned operations should be performed using route and cell details as the key.
+In order to support all these functions with O(log n) time complexity we have developed a data structure called \textit{NodeList}.
+The remainder of this section describes how we have achieved this time complexity with the \textit{NodeList} data structure and how we implemented queries using that data structure.
 
-\subsection{NodeList data structure}
+\subsection{NodeList Data Structure}
 
@@
 Operation Signature & Time Complexity \\ \hline \hline
-add(key : Object, value : NodeValue) & O(log(n)) \\ \hline 
+add(key : Object, value : NodeValue) & O(log n) \\ \hline 
 get(key : Object) : NodeValue & O(1) \\ \hline
-remove(key : Object) : NodeValue & O(log(n)) \\ \hline
-decrementPosition(key : Object) & O(log(n)) \\ \hline
-incrementPosition(key : Object) & O(log(n)) \\ \hline
+remove(key : Object) : NodeValue & O(log n) \\ \hline
+decrementPosition(key : Object) & O(log n) \\ \hline
+incrementPosition(key : Object) & O(log n) \\ \hline
 getTopValues() : List<NodeValue> & O(1) \\ \hline
@@
 \texttt{GetTopValues} operation returns the top ten values of the list.
-The query algorithm sections further elaborate usage of these methods.
+The query algorithm sections further elaborate on the usage of these methods.
 
 One way of implementing this data structure is to use a heap.
-A heap supports adding and extracting maximum value operations in O(log(n)) time.
-However since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back.
-On the other hand if we use a doubly linked list to keep values, it will take O(n) time for  \texttt{decrementPostion} and \texttt{incrementPosition} operations.
+A heap supports adding and extracting maximum value operations in O(log n) time.
+However, since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back.
+On the other hand if we use a doubly linked list to store values, it will take O(n) time for  \texttt{decrementPostion} and \texttt{incrementPosition} operations.
 Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.
@@
 
-\subsection{OrderedList data structure}
+\subsection{OrderedList Data Structure}
 
@@
 Table \ref{orderedlist_api} shows the operations of the \textit{OrderedList} data structure.
-First six methods has the same meaning as the \textit{NodeList} data structure and last two can be used to get the size of the list and the value of the last object.
+First six methods have the same meaning as the \textit{NodeList} data structure and last two can be used to get the size of the list and the value of the last object.
 We internally keep a counter and a pointer to tail to support these operations in O(1) time.
@@
         \includegraphics[width=3.0in]{orderedlist.png}
-        \caption{\textit{OrderedList} data structure implementation}
+        \caption {A schematic representation of the \textit{OrderedList} data structure}
         \label{orderedlist_impl}
@@
 
-\subsection{DynamicHeap data structure}
+\subsection{DynamicHeap Data Structure}
 
@@
 Operation Signature & Time Complexity \\ \hline \hline
-add(key : Object, value : NodeValue) & O(log(n)) \\ \hline
-remove(key : Object) : NodeValue & O(log(n)) \\ \hline
+add(key : Object, value : NodeValue) & O(log n) \\ \hline
+remove(key : Object) : NodeValue & O(log n) \\ \hline
 containsKey(key : Object) : boolean & O(1) \\ \hline
 get(key : Object) : NodeValue & O(1) \\ \hline
-moveUp(key : Object) & O(log(n)) \\ \hline
-moveDown(key : Object) & O(log(n)) \\ \hline
+moveUp(key : Object) & O(log n) \\ \hline
+moveDown(key : Object) & O(log n) \\ \hline
 getMaxkey() : Object & O(1) \\ \hline
-extractMax() : NodeValue & O(log(n)) \\ \hline
+extractMax() : NodeValue & O(log n) \\ \hline
 \end{tabular}
@@
 As shown in figure \ref{dynamicheap_impl} \textit{HeapNode} object contains the array index and the \textit{NodeValue} given by the user.
-This allows us to retrieve array index in O(1) time and move the element O(log(n)) time.
+This allows us to retrieve array index in O(1) time and move the element O(log n) time.
 
@@
         \includegraphics[width=3.0in]{DinamicHeap.png}
-        \caption{\textit{DynamicHeap} data structure implementation}
+        \caption {A schematic representation of the \textit{DynamicHeap} data structure}
         \label{dynamicheap_impl}
@@
 
-\subsection{Finding Runing Median}
-In general, we can find runing median using two heaps.
+\subsection{Finding the Runing Median}
+In general, we can find the runing median using two heaps.
 One to keep values greater than the current median and one to keep values less than the current median.
-An implement of this algorithm using two \textit{java.util.PriorityQueue}s have an O(log(n)) time complexity for add operation and O(n) time complexity for remove operation (\texttt{remove} operation of \textit{java.util.PriorityQueue} takes O(n) time).
+An implementation of this algorithm using two \textit{java.util.PriorityQueue}s have an O(log n) time complexity for add operation and O(n) time complexity for remove operation (\texttt{remove} operation of \textit{java.util.PriorityQueue} takes O(n) time).
 We overcame this issue using \textit{DynamicHeap} to keep values.
-In our solution, we use value as the key to insert and delete from the \textit{DynamicHeap} and keep value with its count (same value can be repeated) in the \textit{DynamicHeap}.
-In worst case it has O(log(n)) time complexity for both add and remove operations and in best case (i.e value already present and that does not cause median change) it has O(1) time complexity.
+In our solution, we use value as the key to insert and delete from the \textit{DynamicHeap} and keep each value with its count (same value can be repeated) in the \textit{DynamicHeap}.
+In the worst case this has a O(log n) time complexity for both the add and remove operations and in the best case (i.e value is already present and it does not cause a median change) it has O(1) time complexity.
 
@@
 Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.
-Since queue operations are supported in O(1) time, query evaluations happen in O(log(n)) time.
+Since queue operations are supported in O(1) time, query evaluations happen in O(log n) time.
 Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to evaluate the query an event.
@@
 \begin{enumerate}
-	\item The profitability is calculated using the median profit and the number of empty taxis.
-In order to calculate median, we use \textit{java.util.PriorityQueue} based approach mentioned in (Section 2.4).
+	\item Profitability is calculated using the median profit and the number of empty taxis.
+In order to calculate the median, we use \textit{java.util.PriorityQueue} based approach mentioned in (Section 2.4).
 We choose this option over our implementation, because this approach performs better with smaller window sizes.
 First, these heap sizes do not grow over 145 and there are same values repeating in the list increasing the probability of finding a value without traversing the whole list.
-Secondly, our implementation creates more objects per operation compared to just keeping double values though the time complexity is O(log(n)).
+Secondly, our implementation creates more objects per operation compared to just keeping double values though the time complexity is O(log n).
 In section 4.3 we show that our approach works better for larger window sizes.
@@
 One way of doing this is to send all events for the same process instance and process both details with one event.
-However this avoids the possibility of partitioning data across different processes.
+However, this avoids the possibility of partitioning data across different processes.
 We generate two events, one for pickup data and one for dropoff data to make the process parallel.

### 2015-05-07 paper/experiments.tex
making the camara copy

@@
 
-Figure \ref{routegraph} shows the process graph used to process frequent route queries.
+Figure \ref{routegraph} shows the process graph used to process frequent routes queries.
 \textit{RouteEventEmitter} reads the data file and emits events with a sequence number.
@@
 This problem can  possibly fixed by adding a limit to the message queue size.
-On the other hand in a real time system, this event ordering many not required since event processing happens at real time.
+On the other hand in a real-time system, this event ordering many not required since event processing happens at real-time.
 

### 2015-05-07 paper/introduction.tex
making the camara copy

@@
 
-If we consider an application that must process the 173 million trips dataset within an year, it would be a trivial task and several simple implementations would satisfy that requirement.
+If we consider an application that must process the 173 million trips dataset within an year, several simple implementations would satisfy that requirement.
 However if we consider an application that must process these data within minutes with the added constraint of minimal latency to process each event, we observe following research challenges.
 \begin{enumerate}
-	\item \textit{Handling thousands of values in real time :}  An algorithm with O(n) time complexity will not scale for thousands. How can we design an algorithm with O(log n) time complexity?
+	\item \textit{Handling thousands of values in real-time :}  An algorithm with O(n) time complexity will not scale for thousands. How can we design an algorithm with O(log n) time complexity?
 	\item \textit{Parallel evaluation :}  This application does not inherently support parallel evaluation.
@@
 \subsection{Contributions}
-We observe following key contributions, which can be used in solving future DEBS challenges as well.
+We observe following key contributions, which can be used in solving future DEBS grand challenges as well.
 \begin{enumerate}
@@
 	\item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity.
-One of the key challenges of DEBS 2014 \cite{jerzak2014debs}  was to find the median of a large time window.
+One of the key challenges of 2014 problem \cite{jerzak2014debs}  was to find the median of a large time window.
 We believe our solution is useful in such situations.

### 2015-05-07 paper/query_evaluation.tex
making the camara copy

@@
 
-In this section, we explain how to evaluate queries with the given data set.
-For this query evaluation we can use a distributed event stream processing system with the process graph as shown in Figure \ref{twonode_graph} for both queries.
+In this section, we explain how to evaluate queries with the given dataset.
+For this query evaluation, we can use a distributed event stream processing system with the process graph as shown in Figure \ref{twonode_graph} for both queries.
 
@@
 \begin{algorithm}
-\caption{Algorithm to aggregate top ten results from QueryEvaluators and generate global events}
+\caption{Algorithm to aggregate top ten results from \textit{QueryEvaluators} and generate global events}
 \label{aggregate_algorithm}

### 2015-05-07 paper/query_implementation.tex
making the camara copy

@@
 get(key : Object) : NodeValue & O(1) \\ \hline
+containsKey(key : Object) : boolean & O(1) \\ \hline
 remove(key : Object) : NodeValue & O(log n) \\ \hline
@@
 Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.
-A \textit{Map} which points to values can be used to retrieve values using a key.
 
@@
 
-\subsection{Frequent Route Query}
+\subsection{Frequent Routes Query}
 
-The frequent route query entails finding the most frequent routes within the last 30 minutes.
+The frequent routes query entails finding the most frequent routes within the last 30 minutes.
 Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.
 Since queue operations are supported in O(1) time, query evaluations happen in O(log n) time.
-Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to evaluate the query an event.
+Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to evaluate the query for an event.
 This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an object called \textit{RouteCount}.

### 2015-05-07 paper/related_work.tex
making the camara copy

@@
 \section{Related work}
-DEBS Grand Challenge problem naturally fallen into distributed stream processing domain.
+DEBS Grand Challenge problem naturally fall into the distributed stream processing domain.
 Many such systems have been developed with different feature sets according to the applications they use.
-Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track trends in web queries.
+Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click-through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track trends in web queries.
 All these systems are based on the Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data.
-Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence and dynamic membership management.
+Apart from basic communication these frameworks provide other features such as reliable message delivery, state persistence and dynamic membership management.
+Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets \cite{zaharia2012resilient}.
 
 Several approaches has been developed for DEBS Grand Challenge 2014 \cite{jerzak2014debs}.
-Predictive Load Management in Smart Grid Environments \cite{mutschler2014predictive} provides a solutions based on EventCore.
-They have developed an efficient algorithms to find median.
+Predictive Load Management in Smart Grid Environments \cite{mutschler2014predictive} provides a solution based on EventCore.
+That involved development of efficient algorithms to find median.
 Scalable Stateful Stream Processing for Samart Grids \cite{fernandez2014scalable} implemented their solution on SEEP \cite{castro2013integrating} a scalable fault tolerant framework.

### 2015-05-07 paper/suriarachchi.tex
making the camara copy

--- 2015-04-28 paper/paper.tex
@@
 
-\documentclass{acm_proc_article-sp}
+%\documentclass{acm_proc_article-sp}
+\documentclass{sig-alternate}
 
@@
 
-\title{A High-throughput, Scalable Solution for the New York Taxi Trips DEBS Grand Challenge Problem}
+\newfont{\mycrnotice}{ptmr8t at 7pt}
+\newfont{\myconfname}{ptmri8t at 7pt}
+\let\crnotice\mycrnotice%
+\let\confname\myconfname%
+
+\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
+Copyrights for components of this work owned by others than ACM must be honored.
+Abstracting with credit is permitted.
+To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.
+Request permissions from Permissions@acm.org.}
+\conferenceinfo{DEBS$^,$ 15,} {June 29 - July 3, 2015, OSLO, Norway.}
+\copyrightetc{Copyright 2015 ACM \the\acmcopyr}
+\crdata{978-1-4503-3286-6/15/06\ ...\$15.00.\\
+http://dx.doi.org/10.1145/2611286.2611330.}
+
+\title{DEBS Grand Challenge: A High-Throughput, Scalable Solution for Calculating Frequent Routes and Profitability of New York Taxis}
 %
@@
 \begin{abstract}
-Processing complex queries on unbounded event streams in real time, is a challenge for many data processing systems.
-These systems are expected to process data with reduced latency to generate real time-events, and at high throughput to minimize the required hardware.
-In this regard, DEBS Grand Challenge 2015 focuses on evaluating two queries (frequent routes and  profitable cells) in real time with low latency and high throughput.
+Processing complex queries on unbounded event streams in real-time, is a challenge for many data processing systems.
+These systems are expected to process data with reduced latency to generate real-time events, and at high throughput to minimize the required hardware.
+In this regard, DEBS Grand Challenge 2015 focuses on evaluating two queries (frequent routes and  profitable cells) in real-time with low latency and high throughput.
 These queries involve processing windows of thousands of records.
@@
 In this paper, we present a set of data structures that we designed to evaluate the aforementioned queries with O(log n) time complexity and a data partitioning technique to evaluate them in parallel.
-We then evaluate our solution on a signal machine as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN.
-We are able to process the frequent routes query with the 173 million trips dataset within 5 minutes with less than 4 millisecond latency and the profitable cells query with same dataset within 11 minutes with less than 5 millisecond latency.
+We then evaluate our solution on a single machine as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN.
+We were able to process the frequent routes query with the 173 million trips dataset within 5 minutes with less than 4 millisecond latency and the profitable cells query with same dataset within 11 minutes with less than 5 millisecond latency.
 \end{abstract}
@@
 % A category with the (minimum) three required fields
+\category{C.2.4}{Distributed Systems}{Distributed applications}
 %A category including the fourth, optional field follows...
-
-\keywords{DEBS Grand Challenge, Distributed Event Processing, Time Complexity} % NOT required for Proceedings
+\keywords{Distributed Event Processing, Real-time analytics, Running median, Time Complexity, Event Stream Processing} % NOT required for Proceedings
 
@@
 \input{query_implementation}
-\clearpage
 \input{query_evaluation}
@@
 \input{conclustion}
+\newpage
 %

### 2015-05-13 presentation/grand_challenge.tex
adding the presentation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<handout>
{
  \usepackage{pgf}
  \usepackage{pgfpages}

\pgfpagesdeclarelayout{4 on 1 boxed}
{
  \edef\pgfpageoptionheight{\the\paperheight} 
  \edef\pgfpageoptionwidth{\the\paperwidth}
  \edef\pgfpageoptionborder{0pt}
}
{
  \pgfpagesphysicalpageoptions
  {%
    logical pages=4,%
    physical height=\pgfpageoptionheight,%
    physical width=\pgfpageoptionwidth%
  }
  \pgfpageslogicalpageoptions{1}
  {%
    border code=\pgfsetlinewidth{2pt}\pgfstroke,%
    border shrink=\pgfpageoptionborder,%
    resized width=.5\pgfphysicalwidth,%
    resized height=.5\pgfphysicalheight,%
    center=\pgfpoint{.25\pgfphysicalwidth}{.75\pgfphysicalheight}%
  }%
  \pgfpageslogicalpageoptions{2}
  {%
    border code=\pgfsetlinewidth{2pt}\pgfstroke,%
    border shrink=\pgfpageoptionborder,%
    resized width=.5\pgfphysicalwidth,%
    resized height=.5\pgfphysicalheight,%
    center=\pgfpoint{.75\pgfphysicalwidth}{.75\pgfphysicalheight}%
  }%
  \pgfpageslogicalpageoptions{3}
  {%
    border code=\pgfsetlinewidth{2pt}\pgfstroke,%
    border shrink=\pgfpageoptionborder,%
    resized width=.5\pgfphysicalwidth,%
    resized height=.5\pgfphysicalheight,%
    center=\pgfpoint{.25\pgfphysicalwidth}{.25\pgfphysicalheight}%
  }%
  \pgfpageslogicalpageoptions{4}
  {%
    border code=\pgfsetlinewidth{2pt}\pgfstroke,%
    border shrink=\pgfpageoptionborder,%
    resized width=.5\pgfphysicalwidth,%
    resized height=.5\pgfphysicalheight,%
    center=\pgfpoint{.75\pgfphysicalwidth}{.25\pgfphysicalheight}%
  }%
}

  \pgfpagesuselayout{4 on 1 boxed}[a4paper, border shrink=5mm, landscape]
  \nofiles
}

\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

%\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage[vlined]{algorithm2e}
\usepackage{caption}
\usepackage{subcaption}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Grand Challenge 2015]{DEBS Grand Challenge: A High-Throughput, Scalable Solution for Calculating Frequent Routes and Profitability of New York Taxis} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Amila Suriarachchi} % Your name
\institute[CSU] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Colorado State University \\ % Your institution for the title page
\medskip
\textit{amilas@cs.colostate.edu} % Your email address
\medskip
}
\date{May 4, 2015} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\frametitle{Outline} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{Introduction} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
%------------------------------------------------

\begin{frame}
\frametitle{Problem Characteristics}
\begin{itemize}
	\item The whole dataset contains 173 million records within an year
	\item There are 899010 unique routes and 34 unique cells
	\item 30 minute time window contains around 10000 events assuming equal event distribution
		\begin{itemize}
			\item Thousands of route and cell details need to process with each event
		\end{itemize}
	\item Top ten frequent routes and profitable cells should be calculated whole dataset
		\begin{itemize}
                        \item Dataset can not be simply partitioned 
                \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Research Challenges}
\begin{itemize}
	\item Handling thousands of values in real-time
		\begin{itemize}
			\item An algorithm with O(n) time complexity will not scale for thousands. How can we design an algorithm with O(log n) time complexity?
		\end{itemize}
	\item Parallel evaluation
		\begin{itemize}
                        \item Partitioning route sets and cell sets triggers events for subsets, but application requires the events with whole sets. Therefore is it possible to partition data and still generate required results to support parallelism? 
                \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Contributions}
\begin{itemize}
        \item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity
        \item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity 
	\item We have developed a scalable parallel evaluation technique to process queries efficiently
\end{itemize}

\end{frame}

\section{Query Implementation}

\begin{frame}
\frametitle{NodeList Data Structure}

\begin{table}
\centering
\begin{tabular}{|l|l|} \hline
Operation Signature & Time Complexity \\ \hline \hline
add(key : Object, value : NodeValue) & O(log n) \\ \hline
get(key : Object) : NodeValue & O(1) \\ \hline
containsKey(key : Object) : boolean & O(1) \\ \hline
remove(key : Object) : NodeValue & O(log n) \\ \hline
decrementPosition(key : Object) & O(log n) \\ \hline
incrementPosition(key : Object) & O(log n) \\ \hline
getTopValues() : List \textless NodeValue \textgreater & O(1) \\ \hline
\end{tabular}
\end{table}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{nodeliststruct.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{OrderedList Data Structure}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{orderedlist.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{OrderedList Data Structure}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{orderedlist.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{DynamicHeap Data Structure}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{DinamicHeap.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Runing Median (1/2)}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{runningmediang.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Runing Median (2/2)}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{runningmediand.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Frequent Routes Query}

\begin{algorithm}[H]
\footnotesize
before $\gets$ nodeList.getTopValues()\;
window.add(event)\;
\eIf{ nodeList.containsKey(event.route) }{
        routCount $\gets$ nodeList.get(event.route)\;
        routCount.count++\;
        nodeList.decrementPosition(event.route)\;
}{
        nodeList.add(event.route, new RouteCount(1))\;
}

\While{ there exists expired events }{
        expEvent $\gets$ window.poll()\;
        routCount $\gets$ nodeList.get(expEvent.route)\;
        routCount.count--\;
        \eIf{ routCount.count is 0 }{
                nodelist.remove(expEvent.route)\;
        }{
                nodeList.incrementPosition(expEvent.route)\;
        }
}

now $\gets$ nodeList.getTopValues()\;

\If{ before not equals now }{
        create new top ten route event\;
}

\end{algorithm}

\end{frame}

\section{Query evaluation}

\begin{frame}
\frametitle{Sequential Evaluation}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{sequential.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Parallel Evaluation}

\begin{figure}
        \includegraphics[width=0.8\linewidth]{parallel.png}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Result Aggregator}

\begin{algorithm}[H]
\footnotesize
before $\gets$ nodeList.getTopValues()\;
\For {key in event.removedKeys}{
        nodeList.remove(key)\;
}

\For {nodeValue in event.newValuesList}{
        \eIf {nodeList.containsKey(nodeValue.key)}{
                existingValue $\gets$ nodeList.get(nodeValue.key)\;
                \eIf { existing value is lesser than current }{
                        update the existing value\;
                        nodeList.decrementPosition(nodeValue.key)\;
                }{
                        update the existing value\;
                        nodeList.incrementPosition(nodeValue.key)\;
                }
        }{
                nodeList.add(nodeValue.key, nodeValue)\;
        }
}

now $\gets$ nodeList.getTopValues()\;
\If { before not equals now }{
        emit new event\;
}

\end{algorithm}

\end{frame}

\section{Experiments}

\begin{frame}
\frametitle{Frequent Routes Query}

\begin{figure}
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{throughput_route.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{latency_route.png}
        \end{subfigure}
\end{figure}

\footnotesize Initially throughput increases with the number of route process instances and comes to a steady state. For distributed setup latency comes to a steady state with higher process instances.

\end{frame}

\begin{frame}
\frametitle{Profitable Cells Query}

\begin{figure}
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{throughput_profit.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\textwidth}
                \includegraphics[width=\textwidth]{latency_profit.png}
        \end{subfigure}
\end{figure}

\footnotesize Initially throughput increases with the number of route process instances and comes to a steady state. For distributed setup latency comes to a steady state with higher process instances.

\end{frame}

\end{document}

### 2015-05-13 presentation/grand_challenge.tex
modify the presentation

@@
 
-\author{Amila Suriarachchi} % Your name
+\author[Amila Suriarachchi]{Amila Suriarachchi \\ Shrideep Pallickara } % Your name
 \institute[CSU] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
@@
 }
-\date{May 4, 2015} % Date, can be changed to a custom date
+\date{\today} % Date, can be changed to a custom date
 
@@
 \begin{frame}
-\frametitle{OrderedList Data Structure}
-
-\begin{figure}
-        \includegraphics[width=0.8\linewidth]{orderedlist.png}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}
 \frametitle{DynamicHeap Data Structure}
@@
 \begin{figure}
-        \includegraphics[width=0.8\linewidth]{DinamicHeap.png}
+        \includegraphics[width=0.7\linewidth]{DinamicHeap.png}
 \end{figure}
@@
 
-\footnotesize Initially throughput increases with the number of route process instances and comes to a steady state.
-For distributed setup latency comes to a steady state with higher process instances.
+\footnotesize For distributed setup throughput increases up to 4 processors while latency remain in a constant after 5 processors.
+For single JVM setup throughput increases up to 2 nodes and slightly decreases with the number of processors while latency slightly increases with the number of processors.
 
@@
 
-\footnotesize Initially throughput increases with the number of route process instances and comes to a steady state.
-For distributed setup latency comes to a steady state with higher process instances.
+\footnotesize For distributed set up throughput linearly increases and latency increases with higher number of calculators.
+For single JVM set up throughput linearly increases up to 3 instances and latency slightly increases with the number of calculators.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Scalability with Window Size}
+
+\begin{figure}
+        \includegraphics[width=0.6\linewidth]{window.png}
+\end{figure}
+
+\footnotesize Both implementations scale even for 64h windows. DynamicHeap based algorithm performs better at higher window sizes.
+
+\end{frame}
+
+\section{Conclusion}
+
+\begin{frame}
+\frametitle{Conclusion}
+\begin{itemize}
+        \item Our queries execute with O(log n) time complexity and hence scale with window size.
+        \item Our partitioning technique can be used to scale up the system while producing the desired results. 
+	\item We propose an algorithm to find the running median with O(log n) time complexity.
+\end{itemize}
 

### 2015-05-13 presentation/grand_challenge.tex
fixed version

@@
 
-\documentclass{beamer}
+\documentclass[]{beamer}
 
@@
 \begin{itemize}
-	\item The whole dataset contains 173 million records within an year
-	\item There are 899010 unique routes and 34 unique cells
+	\item The whole dataset contains 173 million records within year 2013
+	\item There are 899010 unique routes and 56697 unique cells
 	\item 30 minute time window contains around 10000 events assuming equal event distribution
@@
 
+\begin{frame}
+\frametitle{Profitable Cells Query}
+\begin{itemize}
+        \item Median fare calculation
+		\begin{itemize}
+			\item Use the PriorityQueue based running median algorithm
+		\end{itemize}
+        \item Handling empty taxis
+		\begin{itemize}
+                        \item Generate two events (Pickup event and Drop event) for each dataset event
+                \end{itemize}
+        \item Handling drop event expiration
+		\begin{itemize}
+                        \item Keep taxi ids within the profitable cell object
+                \end{itemize}
+\end{itemize}
+
+\end{frame}
+
 \section{Query evaluation}

### 2015-05-19 presentation/grand_challenge.tex
updating the presentation

@@
 
-\documentclass[]{beamer}
+\documentclass[handout]{beamer}
 
@@
 \medskip
-\textit{amilas@cs.colostate.edu} % Your email address
+\textit{\{amilas,shrideep\}@cs.colostate.edu} % Your email address
 \medskip
@@
 
-\begin{frame}
-\frametitle{Contributions}
-\begin{itemize}
-        \item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity
-        \item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity 
-	\item We have developed a scalable parallel evaluation technique to process queries efficiently
-\end{itemize}
-
-\end{frame}
+%\begin{frame}
+%\frametitle{Contributions}
+%\begin{itemize}
+%        \item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity
+%        \item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity 
+%	\item We have developed a scalable parallel evaluation technique to process queries efficiently
+%\end{itemize}
+%\end{frame}
 
@@
 \begin{frame}
-\frametitle{NodeList Data Structure}
+\frametitle{NodeList Data Structure (1/2)}
+
+\begin{itemize}
+        \item Provides notion of a sorted list of values starting from position 0
+        \item Values can be accessed using the key (eg. Route, Cell)
+\end{itemize}
 
@@
 
+\end{frame}
+
+\begin{frame}
+\frametitle{NodeList Data Structure (2/2)}
+
+\begin{itemize}
+        \item \textit{OrderedList} data structure is used to keep first ten values
+        \item \textit{DynamicHeap} data structure is used to keep other values
+\end{itemize}
+
 \begin{figure}
@@
 
+\begin{itemize}
+        \item A Doubly linked List is used to store values
+        \item A Map is used to retrieve \textit{ListNode} objects with O(1) time
+\end{itemize}
+
 \begin{figure}
-        \includegraphics[width=0.8\linewidth]{orderedlist.png}
+        \includegraphics[width=0.7\linewidth]{orderedlist.png}
 \end{figure}
@@
 
+\begin{itemize}
+        \item A Heap is used to store values
+        \item \textit{HeapNode} keeps the position of the array to remove, moveUp and moveDown objects with O(log n) time
+	\item A Map is used to retrieve \textit{HeapNode} objects with O(1) time
+\end{itemize}
+
 \begin{figure}
-        \includegraphics[width=0.7\linewidth]{DinamicHeap.png}
+        \includegraphics[width=0.6\linewidth]{DinamicHeap.png}
 \end{figure}
@@
 \begin{frame}
-\frametitle{Runing Median (1/2)}
+\frametitle{Running Median (1/2)}
+
+\begin{itemize}
+        \item Two heaps are used to keep values greater than the current median and less than the current median
+        \item A \textit{PriorityQueue} based implementation
+		\begin{itemize}
+        		\item Takes O(log n) time to add a value 
+        		\item Takes O(n) time to remove a value
+		\end{itemize}
+\end{itemize}
 
 \begin{figure}
-        \includegraphics[width=0.8\linewidth]{runningmediang.png}
+        \includegraphics[width=0.6\linewidth]{runningmediang.png}
 \end{figure}
@@
 \begin{frame}
-\frametitle{Runing Median (2/2)}
+\frametitle{Running Median (2/2)}
+
+\begin{itemize}
+        \item Two heaps are used to keep values greater than the current median and less than the current median
+        \item A \textit{DynamicHeap} based implementation with value as the heap key and count as the heap value 
+                \begin{itemize}
+                        \item Takes O(log n) time to add and remove a value
+                \end{itemize}
+\end{itemize}
 
 \begin{figure}
-        \includegraphics[width=0.8\linewidth]{runningmediand.png}
+        \includegraphics[width=0.6\linewidth]{runningmediand.png}
 \end{figure}
@@
 \eIf{ nodeList.containsKey(event.route) }{
-        routCount $\gets$ nodeList.get(event.route)\;
-        routCount.count++\;
+        routeCount $\gets$ nodeList.get(event.route)\;
+        routeCount.count++\;
         nodeList.decrementPosition(event.route)\;
@@
         expEvent $\gets$ window.poll()\;
-        routCount $\gets$ nodeList.get(expEvent.route)\;
-        routCount.count--\;
-        \eIf{ routCount.count is 0 }{
+        routeCount $\gets$ nodeList.get(expEvent.route)\;
+        routeCount.count--\;
+        \eIf{ routeCount.count is 0 }{
                 nodelist.remove(expEvent.route)\;
@@
 		\begin{itemize}
-			\item Use the PriorityQueue based running median algorithm
+			\item Use the \textit{PriorityQueue} based running median algorithm
 		\end{itemize}
@@
 
+\begin{itemize}
+        \item One \textit{QueryEvaluator} instance is used
+        \item \textit{EventEmitter} and \textit{QueryEvaluator} can be executed in,
+                \begin{itemize}
+                        \item A single JVM using one thread
+                        \item A single JVM using different threads with queues to communicate
+                        \item A distributed setup with TCP to communicate 
+                \end{itemize}
+\end{itemize}
+
 \begin{figure}
-        \includegraphics[width=0.8\linewidth]{sequential.png}
+        \includegraphics[width=0.7\linewidth]{sequential.png}
 \end{figure}
@@
 \frametitle{Parallel Evaluation}
+
+\begin{itemize}
+        \item Multiple \textit{QueryEvaluator} instances are used
+        \item \textit{EventEmitter} and \textit{QueryEvaluators} can be executed in,
+                \begin{itemize}
+                        \item A Single JVM using different threads with queues to communicate
+                        \item A distributed setup with TCP to communicate
+                \end{itemize}
+\end{itemize}
 
